---
title: "0.2_explore-external-data"
author: "Devi Veytia"
date: "2024-04-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
library(dplyr)
library(ggplot2)
```

# Marine energy

```{r load from databases and deduplicate}
# df1 <- readxl::read_excel(here::here("data/raw-data/external/PRIMRE-marine-energy-projects-database.xlsx"),sheet = "Projects")
# df2 <- readxl::read_excel(here::here("data/raw-data/external/Tethys-pnnl-marine-energy-metadata.xlsx"),sheet = "Projects")
# df <- rbind(df1, df2)
# 
# # Deduplicate 
# dups <- revtools::find_duplicates(data=df, match_variable = "project_name", to_lower = TRUE)
# marine_energy_df <- revtools::extract_unique_references(df, dups)
# 
# # save
# save(marine_energy_df, file = here::here("data/raw-data/external/PRIMRE-Tethys-combined.RData"))

load(here::here("data/raw-data/external/PRIMRE-Tethys-combined.RData"))

# Quick plot 
marine_energy_df <- marine_energy_df %>%
  mutate(year = as.numeric(ifelse(is.na(operational_start_year), project_start_year, operational_start_year)),
         max_rated_power_capacity_MW = as.numeric(max_rated_power_capacity_MW)) %>%
  filter(!is.na(year))
marine_energy_df$year <- as.Date(paste0(marine_energy_df$year,"-01-01"))
marine_energy_df <- marine_energy_df %>%
  tidyr::replace_na(list(max_rated_power_capacity_MW = 0)) %>%  # Replace NA with zeros
  group_by(project_life_cycle, year) %>%
  arrange(project_life_cycle, year) %>%  # Arrange by year so adding works
  group_by(project_life_cycle) %>%
  mutate(cumMW = cumsum(max_rated_power_capacity_MW))

summary(marine_energy_df)
  

ggplot(marine_energy_df, aes(year, cumMW, fill = project_life_cycle, col = project_life_cycle))+
  geom_area(alpha = 0.5)+
  scale_x_date()+
  theme_bw()
```

```{r map countries of MRE locations? -- need to do}

```

# IRENA renewable electricity statistics

```{r}
df <- readxl::read_excel(here::here("data/raw-data/external/IRENA-electricity-statistics-by-country-year.xlsx"),
                         sheet = "ELECSTAT-C", col_names = FALSE, skip = 2)
colnames(df) <- c("country","technology","data_type","grid_connection","year","value")

df <- df %>% 
  tidyr::fill(country, technology, data_type, data_type, grid_connection, .direction = "down")

# filter data to one variable, all grid connections
df <- df %>%
  filter(data_type == "Electricity Installed Capacity (MW)", grid_connection == "All", technology %in% c("Marine energy","Offshore wind energy"))

# Format columns
df$year <- as.Date(paste0(df$year,"-01-01"))

# Quick plot
ggplot(df, aes(year, value, fill = country, col = country))+
  facet_grid(technology ~., scales = 'free_y') +
  geom_area(alpha=0.5)+
  labs(y = "Electricity Installed Capacity (MW)")+
  scale_x_date()+
  theme_bw()+
  theme(legend.position = 'none')

# can colour this by country type? 

```



# Novel CDR deployments - State of CDR Report 2022

```{r load in novel CDR deployment data}
cdr_deployment <- readxl::read_excel(here::here("data/raw-data/external/State-of-Carbon-Report-2022-Novel-CDR-Deployment-Data.xlsx"), skip = 1)

carbon_sinks <- unique(do.call(c, unique(as.vector(cdr_deployment[,3:6]))))
carbon_sinks

cdr_deployment$carbon_sink <- apply(cdr_deployment[,3:6], 1, paste, collapse=", ")
cdr_deployment <- cdr_deployment %>%
  mutate(ocean = ifelse(
    grepl("Aquatic", carbon_sink) | grepl("Deep Ocean", carbon_sink)| 
      grepl("Dissolved CO2", carbon_sink) | grepl("Bicarbonate", carbon_sink), TRUE,FALSE
  ))
cdr_deployment$year = as.Date(paste0(cdr_deployment$`Year active`,"-01-01")) 

summary(cdr_deployment)

cdr_deployment %>%
  filter(ocean) %>%
  group_by(year) %>%
  summarise(`Gross Removals\n(tCO2/year)` = sum(`Gross Removals\n(tCO2/year)`, na.rm=T)) %>%
  mutate(`Gross Removals\n(tCO2/year)` = cumsum(`Gross Removals\n(tCO2/year)`)) %>%
  ggplot(aes(year, `Gross Removals\n(tCO2/year)`))+
  geom_line()+
  geom_point()+
  scale_x_date()+
  labs(y = "Cumulative Gross Removals\n(tCO2/year)",
       caption = "Source: State of CDR Report 2022 - Novel CDR Deployment Data\nfiltered for where carbon is stored in ocean sinks")+
  theme_bw()
```


# Increasing efficiency

## Fishing

Estimates of fishing effort were sourced from a global database based on estimates of total vessel engine size and number of fishing days in a year, assembled from FAO, the European Union, regional tuna-management bodies and other sources50,51

Fuel consumption rates were extracted from FEUD19 (litres per round weight tonne). Records of fisheries operating before 1985 were excluded from analysis, as were any records for which target species group or gear type could not be determined

The combined rate of emissions was 3.3 kg CO2-eq GHG per litre of fuel combusted.

## Transport

```{r load in iea total shipping emissions}
shipping_emissions <- readxl::read_excel(here::here("data/raw-data/external/IEA-Energy-End-uses-and-Efficiency-Indicators-database-Highlights-November-2023.xlsx"), sheet = "Transport - Emissions")

colnames(shipping_emissions) <- c("country","vehicle_type","product","2000","2005","2010","2015","2016","2017","2018","2019","2020","2021")

shipping_emissions <- shipping_emissions %>%
  filter(vehicle_type == "Total ships")

shipping_emissions <- reshape2::melt(shipping_emissions, 
                                     measure.vars = c("2000","2005","2010","2015","2016","2017","2018","2019","2020","2021"),
                                     variable.name = "year",
                                     value.name = "value")

shipping_emissions$year <- as.Date(paste0(shipping_emissions$year,"-01-01"))
shipping_emissions$value <- as.numeric(shipping_emissions$value)

shipping_emissions %>%
 filter(country == "IEA Total") %>%
  ggplot(aes(year, value))+
  geom_line()+
  geom_point()+
  scale_x_date()+
  labs(y = "IEA total shipping emissions (Mt CO2)")+
  theme_bw()+
  theme(legend.position = 'none')

```

```{r load in IEA shipping energy efficiency}

shipping_efficiency <- readxl::read_excel(here::here("data/raw-data/external/IEA-shipping-energy-efficiency.xlsx"), sheet = "Data")
shipping_efficiency$year <- as.Date(paste0(shipping_efficiency$year,"-01-01"))

ggplot(shipping_efficiency, aes(year, carbon_intensity_gCO2_per_tkm))+
  geom_line()+
  geom_point()+
  scale_x_date()+
  labs(y="Shipping energy efficiency (gCO2/tkm)", caption = "source: IEA energy end uses and efficiency indicators database")+
  theme_bw()
```







## Shipping








