---
title: "0.8_sample-non-MRE-located-ocean-further-screening"
author: "Devi Veytia"
date: "2025-05-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The problem was that in the oro type screening, not enough of the random sample had instances of other ORO types that were NOT MRE located or ocean. So fit a multilabel classifier for MRE located, MRE ocean or other, and then ran predictions for all unseen mitigation oro branch. I will sample from these for further screening/coding and treat the sample as random because I will only use these articles for selection and prediction.

# Set up

```{r set up libraries}
## Load libraries
library(dplyr)
library(dbplyr)
library(R.utils)
library(RSQLite)
library(ggplot2)
library(tidyverse)

source(here::here("R/bool_detect.R"))
source(here::here("R/write_ris_batches.R"))
```

```{r set the seed}
addTaskCallback(function(...) {set.seed(123);TRUE})
```


```{r Set up inputs to determine how many articles to sample}
setSize <- 200 # write the files in batches to not overwhelm sysrev
#nRecsTotal = 800 # Max no of articles to sample in total per label
nRecsTotal = 2500 # total target number of records to screen
```

# Sample

```{r collect all potential mitigation ORO references}

# Read in all mitigation ORO references 

## connect to database
p1_db <- RSQLite::dbConnect(RSQLite::SQLite(),
                            here::here("data","sqlite-databases","product1.sqlite"),
                            create=FALSE)


# Collect table of unique references metadata
uniquerefs <- tbl(p1_db, "uniquerefs") %>%
  select(analysis_id, title, abstract, keywords) %>%
  collect()

# identify references relevant to mitigation
pred_OROmitigation <- tbl(p1_db, "pred_oro_any_mitigation") %>%  # mitigation 
  select(analysis_id, `oro_any.M_Renewables - mean_prediction`, `oro_any.M_Increase_efficiency - mean_prediction`,
         `oro_any.M_CO2_removal_or_storage - mean_prediction`) %>%
  collect()

# Filter to articles relevant for at least 1 mitigation ORO
pred_OROmitigation <- pred_OROmitigation %>%
  mutate(n_OROs = rowSums(.[2:4] > 0.5)) %>% 
  filter(1 <= n_OROs) %>%
  left_join(uniquerefs, by = "analysis_id") %>%
  filter(!is.na(abstract)) %>%
  rename(id = analysis_id) 
  

# Disconnect from database
dbDisconnect(p1_db)


# Also read in updated references
uniquerefs_update <- read_delim(
  here::here("data/raw-data/unique_references_UPDATE_13-05-2025.txt"), delim = "\t"
)
uniquerefs_update <- uniquerefs_update %>%
  select(analysis_id, title, abstract, keywords)


# Read in predictions of MRE-located, MRE-ocean or other
# Filter to NOT MRE located or ocean
mrePred <- read.csv(
  here::here("outputs/coded-label-predictions/oro_type.MRE_v2_MreLO_predictions.csv")) 


mrePred %>%
  summarise(
    n_MreLocated = sum(0.5 <= `oro_type.MRE.Located...mean_prediction`),
    n_MreOcean = sum(0.5 <= `oro_type.MRE.Ocean...mean_prediction`),
    n_MreAbsent = sum(0.5 <= `oro_type.MRE.Absent...mean_prediction`)
  )
#   n_MreLocated n_MreOcean n_MreAbsent
# 1        17021      13776       29934

mreAbsent <- mrePred %>%
  filter(`oro_type.MRE.Located...mean_prediction` < 0.5 &
           `oro_type.MRE.Ocean...mean_prediction` < 0.5 &
           0.5 <= `oro_type.MRE.Absent...mean_prediction`)

nrow(mreAbsent)  # 29927 (removed articles predicted relevant for multiple)


## Sample these articles and write to .ris files

randids <- sample(mreAbsent$id, round(nRecsTotal))

df <- pred_OROmitigation %>%
  filter(id %in% randids) %>%
  rename(analysis_id = id) %>%
  select(analysis_id, title, abstract, keywords) %>%
  bind_rows(
    uniquerefs_update %>%
      filter(analysis_id %in% randids)
  )
nrow(df)

write_ris_batches(df = df %>% select(title, abstract, keywords), 
                  batchSize = setSize, 
                  writeDir = here::here("data/derived-data/0.8_sample-notMRE-screen-and-code"), 
                  fileNamePrefix = "random-notMRE_")

sampled_articles <- df %>% 
  select(analysis_id, title, abstract, keywords) %>% 
  mutate(sample_method = "random_notMRE") 


# ## Write sampled articles to sqlite
# p1_db <- RSQLite::dbConnect(RSQLite::SQLite(),
#                             here::here("data","sqlite-databases","product1.sqlite"),
#                             create=FALSE)
# dbWriteTable(p1_db, "0.8_sampled-articles_notMRE", sampled_articles, overwrite = FALSE, append = FALSE)
# dbDisconnect(p1_db)
```
