---
title: "2.0_compile-trends-data"
author: "Devi Veytia"
date: "2025-07-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r load libraries}

# general data handing
library(dplyr)
library(dbplyr)
library(R.utils)
library(ggplot2)
library(ggalluvial)
library(tidyr)
library(stringr)
library(viridis)
library(countrycode)
library(broom)
library(conflicted)
library(tidyverse)


conflict_prefer("select", "dplyr")
conflicts_prefer(dplyr::filter)



## AESTHETICS
factor_aes <- readxl::read_excel(here::here("R/mitigation_factor_aesthetics.xlsx"))
typeAES <- factor_aes[which(factor_aes$variable == "oro_type"),]
typeAES <- typeAES[order(typeAES$order),]


```





## Load data

Data structure:
list with each level is 
- data frame for publications: 
- data frame for n policy documents
- data frame for deployment

id variables: oro_type, component (publication, policy, deployment), variable_name,
response variable: y

```{r load data}

load(here::here("data", "derived-data", "mitigationORO_pubs.RData")) #pubs
# load(here::here("data", "derived-data", "mitigationORO_pubsProp.RData"))
load(here::here("data", "derived-data", "n_nonBindPolicy_docs.RData")) # legDat
load(here::here("data", "derived-data", "n_legislation_docs.RData")) # polDat
load(here::here("data/derived-data/mitigationDeployDat.RData")) # allDeployDat
load(here::here("data/derived-data/mitigationPostsDat.RData")) # postsDat

year_lim <- c(2000, 2024)


# bind all data together
allComponentDat_model <- pubs %>%
  bind_rows(polDat) %>%
  bind_rows(legDat) %>%
  bind_rows(allDeployDat) %>%
  bind_rows(postsDat) %>%
  mutate(
    component = replace(component, component == "non-binding policy", "policy"),
    # y = scale(y, center=TRUE, scale=TRUE)
  ) %>%
  filter(year_lim[1] <= year & year <= year_lim[2])


```


# Transfer entropy

Publications: 

* Behrendt et al. 2019. RTransferEntropy—Quantifyinginformationflowbetweendifferent timeseriesusingeffective transfer entropy
* **Amornbunchornvej et al 2021. Variable-lag Granger Causality and Transfer Entropy for Time Series Analysis**

Uses: To detect significance of information flow from X -> Y with a time lag. unlike Granger Causaility, can detect non-linear relationships.

Data Assumptions:

* evenly spaced, no NA (fill NAs with 0 in our time series)
* stationary data: refers to the idea that the statistical properties of a time series do not change over time. More specifically, a stationary time series is one in which the mean, variance, and autocorrelation structure are constant over time (no noticible changing levels, increasing variance). (use Box-Cox transform to make data normal)
* time lag is fixed (except see variable lag-Transfer entropy with bootstrapping, as bootstrapping approach increases the performance of transfer entropy methods in this task - install.packages("VLTimeCausality"))


Pipeline:

* correct for heteroskedasity using a Box-Cox transformation to make your data normal (caret::BoxCoxTrans)
* get an iterable list of the pairwise edges for the different nodes in the time series
* Use VL-TE (bootstrapped) from the VLTTimeCausality package to get TE ratio and p-value (higher the ratio, stronger evidence for causality)
* Plot the TE ratio for each of the edges to determine the evidence for the relationship 

```{r VL-TE analysis, eval = FALSE}
# remotes::install_github("DarkEyes/VLTimeSeriesCausality")
require(VLTimeCausality)

# Define levels to loop through

# Loop 1 -- different oros
oros <- unique(allComponentDat_model$oro_type)

# Loop 2 -- process all possible edges
# Define nodes and edges to process
nodes <- unique(allComponentDat_model$component)
edges <- expand.grid(nodes, nodes)
# edges <- edges[!apply(edges, 1, is.unsorted), ] # if order doesn't matter? if causality is tested in both directions...
edges <- edges[edges[,1] != edges[,2],]



# Function to make BoxCoxTransformation -- make data normal
BCTransform <- function(series){
  BCMod <- caret::BoxCoxTrans(series)
  series_trans <- predict(BCMod, series)
}

try_BCTransform <- function(series){
  tryCatch(
    {BCTransform(series)},
    error = function(e){
      return(series)
    }
  )
}


# Define function to only use BC transform is HeteroSkedasticity present?
correctHeteroskedasticity <- function(series, years){
  lmMod <- lm(series~years)
  isHet <- lmtest::bptest(lmMod)
  if(isHet$p.value <= 0.05){
    series_trans <- BCTransform(series)
    return(series_trans)
  }else{
    return(series)
  }
}
try_correctHeteroskedasticity <- function(series, years){
  tryCatch(
    {correctHeteroskedasticity(series, years)},
    error = function(e){
      return(series)
    }
  )
}



## Loop 1 -- loop through OROs
TE_results <- data.frame()
TE_results_summary <- data.frame()
for(oro in oros){
  # oro <- "MRE-Ocean" # for testing
  
  # Data processing
  # Fillin missing values with a 0 -- so that same years are represented across time series
  # correct for heteroskedasiticy
  oroDat <- allComponentDat_model %>%
    filter(oro_type == oro) %>%
    complete(component, year=year_lim[1]:year_lim[2], 
             fill = list(y=0), explicit= FALSE) %>%
    group_by(component) %>%
    mutate(y_trans = try_BCTransform(y)) %>%
    ungroup()
  # ggplot(oroDat, aes(x=year, y=y_trans))+geom_line()+facet_wrap(vars(component), scales="free_y")
  
  
  
  edgesResults <- data.frame(
    oro_type = rep(oro, nrow(edges)),
    NodeX = edges[,1],
    NodeY = edges[,2],
    TE_XCauseY = rep(NA, nrow(edges)),
    TE_pval = rep(NA, nrow(edges)),
    TE_ratio = rep(NA, nrow(edges))
  )
  
  # Loop through the edges and test for causality
  for(e in 1:nrow(edges)){

    TE_out <- tryCatch(
      {VLTransferEntropy(
      Y= oroDat$y_trans[oroDat$component == edges[e,1]],
      X= oroDat$y_trans[oroDat$component == edges[e,2]],
      maxLag = 5,
      VLflag=TRUE,nboot=100, alpha = 0.05)},
      error = function(e){
        return(NULL)
      }
    )
    if(is.null(TE_out)){
      edgesResults$TE_pval[e] <- NA
      edgesResults$TE_XCauseY[e] <- NA
      edgesResults$TE_ratio[e] <- NA
    }else{
      edgesResults$TE_pval[e] <- TE_out$pval
      edgesResults$TE_XCauseY[e] <- TE_out$XgCsY_trns # TS$X causes TS$Y TRUE/FALSE
      edgesResults$TE_ratio[e] <- TE_out$TEratio
    }
    
    
  }
  
  # summarize results to most causal direction if a two-way effect found
  # ie. When TRUE for both directions, resolve by picking the direction with the highest TE ratio
  edgesResultsSummary <- data.frame()
  uniqueEdges <- edges[!apply(edges, 1, is.unsorted), ]
  for(ue in 1:nrow(uniqueEdges)){
    
    incl = vector(length = nrow(edgesResults))
    for(i in 1:nrow(edgesResults)){
      tmp <- c(edgesResults[i,c("NodeX","NodeY")]) %in% uniqueEdges[ue,]
      incl[i] <- sum(tmp)==2
    }
    tmpDat <- edgesResults[incl,] %>%
      filter(TE_XCauseY) %>%
      arrange(desc(TE_ratio)) %>%
      slice_head(n=1)
    
    edgesResultsSummary <- edgesResultsSummary %>%
      bind_rows(tmpDat)
  }
  
    
  # Bind results
  TE_results <- TE_results %>%
    bind_rows(edgesResults)
  TE_results_summary <- TE_results_summary %>%
    bind_rows(edgesResultsSummary)
  
}

# clean
rm(oroDat, nodes, edges, TE_out, edgesResults)



## Check data
# View(TE_results)
# View(TE_results_summary)
# TE_results %>% filter(!is.na(TE_XCauseY)) %>%filter(TE_XCauseY == TRUE)  %>% View


# save(TE_results, TE_results_summary, file = here::here("data/derived-data/TE_results.RData"))
```




## Synthesizing results: At the oro level

```{r load TE results}

# TE_results, TE_results_summary
load(here::here("data/derived-data/TE_results.RData"))

```

1. Build a Directed Causal Graph per oro_type
🔧 Description:
Convert each oro_type group into a directed graph:

Nodes = time series variables (e.g. policy, public support)

Directed Edges = significant NodeX → NodeY pairs with TE_pval < α (e.g. 0.05)

Edge weight = TE_ratio (strength of causality)

This structure allows visual exploration, network metrics (e.g. in-degree), and cross-type comparisons.
```{r plot Directed Causal Graph of TE results for each oro}

require(igraph)

oroTypes <- unique(TE_results$oro_type)


# par(mfrow=c(5,2))

for(i in oroTypes){
  # Example for one oro_type
  g <- graph_from_data_frame(
    TE_results[TE_results$oro_type == oroTypes[i] & 
                 TE_results$TE_pval < 0.05 & 
                 !is.na(TE_results$TE_ratio) &
                 TE_results$TE_ratio > 0, 
               c("NodeX", "NodeY", "TE_ratio")], 
    directed = TRUE
  )
  plot(g, edge.width = E(g)$TE_ratio, main = paste0(letters[i],") ",oroTypes[i]),)

}

par(mfrow=c(1,1))
```






2. Normalize and Aggregate Causal Strengths Across oro_types


🔧 Description:
Aggregate the TE results across oro_types to build a meta-causal graph:

For each unique NodeX → NodeY, compute:

N significant: how many oro_types had significant causality. This indicates agreement/consistency

Mean TE_ratio: average causal strength across significant cases.



🧠 Why:
To generalize relationships beyond a specific topic and reveal robust, cross-domain causal signals.


Visualize the aggregated meta-network where:

* Edge width = average TE ratio.
* Edge transparency = consistency (e.g. normalized N significant).
* Node size = degree or centrality.

ALSO: Identify subsystems at the meta level (Node colour)
🔧 Description:
Use community detection or hierarchical clustering to identify subsystems (e.g., policy ↔ public interest loop).

🧠 Why:
Find recurring causal motifs or domain-specific differences.


```{r meta network aggregated across oro types}

meta_links <- TE_results %>%
  filter(TE_pval < 0.05,TE_XCauseY==TRUE, !is.na(TE_ratio), TE_ratio > 0) %>%
  group_by(NodeX, NodeY) %>%
  summarise(
    n_sig = n(),
    avg_ratio = mean(TE_ratio, na.rm = TRUE),
    .groups = 'drop'
  )%>%
  mutate(
    relative_nsig = n_sig / max(n_sig)
  ) %>%
  arrange(desc(relative_nsig), desc(avg_ratio)) 


# Make meta graph
meta_graph <- graph_from_data_frame(meta_links, directed = TRUE)

# Calculate clusters
clusters <- cluster_walktrap(meta_graph)


# Visualize meta network
# arrow width ~ average TE_ratio
# arrow colour ~ agreement
# node colour ~ cluster
E(meta_graph)$weight <- meta_links$avg_ratio
E(meta_graph)$color <- scales::alpha("black", meta_links$relative_nsig)

svg(file = here::here("figures/supplemental/meta_TE_network_diagraph.svg"),width = 8, height = 8)
plot(meta_graph, edge.width = log1p(E(meta_graph)$weight) * 1.5,vertex.color = clusters$membership)
dev.off()

```

4. Quantify Node Importance and Influence
🔧 Description:
Use network centrality metrics:

* Out-degree / in-degree: raw count of causal effects given/received.
* Eigenvector or PageRank: influence in the broader network.
* Betweenness: mediators (e.g. public interest as intermediary).

🧠 Why:
Identifies "driver" vs "responder" nodes across domains.
```{r node importance and influence}
# causal influencers
degree(meta_graph, mode = "out") %>% sort(decreasing = T)   
# public support   public interest weighted             publications

# causal responders
degree(meta_graph, mode = "in") %>% sort(decreasing = T)   
# public interest weighted           public support               deployment

# general influence
sort(eigen_centrality(meta_graph)$vector, decreasing =TRUE)  
# public interest public interest weighted              legislation

# mediators
sort(betweenness(meta_graph), decreasing=TRUE)
# deployment           public support             publications

```





Cluster or compare the causal network structure across oro_type groups to identify:

* Similar causal pathways or influence patterns.
* oro_types that share a structural "causal fingerprint".

Method:

* Build a causal graph (adjacency matrix) for each oro_type.
* Vectorize or embed each graph to a comparable format.
* Compute distance or similarity between oro_type graphs.
* Cluster the oro_types.
* Visualize with dendrograms or MDS plots.
```{r compare networks by oro cluster and plot dendogram and pheatmap separately}

## Adjacency matrix

nodes <- unique(c(TE_results$NodeX, TE_results$NodeY))
edge_names <- as.vector(outer(nodes, nodes, FUN = function(x, y) paste0(x, " - ", y)))

# Create adjacency matrix for each oro_type
make_adj_matrix <- function(df, nodes, edge_names) {
  
  
  # adj <- matrix(0, nrow = length(nodes), ncol = length(nodes),
  #               dimnames = list(nodes, nodes))
  adj <- rep(NA, length(edge_names))
  names(adj) <- edge_names
  
  sig_edges <- df %>% filter(TE_pval < 0.05)
  
  for (i in 1:nrow(sig_edges)) {
    
    x <- sig_edges$NodeX[i]
    y <- sig_edges$NodeY[i]
    edge_name <- paste0(x, " - ", y)
    
    if (!is.na(x) && !is.na(y)) {
      adj[edge_name] <- sig_edges$TE_ratio[i]
    }
  }
  return(adj)
}

# One matrix per oro_type
oroTypes <- unique(TE_results$oro_type)
oro_adj_list <- list()
for(oro in oroTypes){
  oro_adj_list[[oro]] <- make_adj_matrix(TE_results %>% filter(oro_type == oro), nodes = nodes, edge_names = edge_names)
}


# oro_adj_list <- TE_results %>%
#   # filter(!is.na(TE_ratio), 0 < TE_ratio, 0.05 <= TE_pval) %>%
#   group_split(oro_type) %>%
#   setNames(unique(TE_results$oro_type)) %>%
#   lapply(make_adj_matrix, nodes = nodes, edge_names = edge_names)


# Vectorize (row-wise)
# Each adjacency matrix is flattened into a vector so we can compute pairwise distances.
adj_vectors <- lapply(oro_adj_list, function(mat) as.vector(t(mat)))
adj_matrix <- do.call(rbind, adj_vectors)
rownames(adj_matrix) <- names(oro_adj_list)
colnames(adj_matrix) <- edge_names

# Compute distance matrix between oro types
# Remove MRE-Bio which makes NA
#adj_matrix[-which(rownames(adj_matrix)=="MRE-Bio"),]
dist_adj_mat <- adj_matrix
dist_adj_mat[is.na(dist_adj_mat)] <- 0
dist_matrix <- stats::dist(dist_adj_mat, method = "euclidean")


## Cluster the oro types
hc <- hclust(dist_matrix, method = "ward.D2")

svg(filename=here::here("figures/supplemental/TE_network_oro_cluster_dendogram.svg"), width=7, height=6)
plot(hc, main = "Dendogram of TE causal networks clustering by ORO type", xlab = "")
dev.off()


adj_matrix_subset <- adj_matrix[,apply(adj_matrix, 2, function(x){max(x, na.rm=T)>0})]
adj_matrix_subset <- adj_matrix_subset[,!is.na(colnames(adj_matrix_subset))]

## Look at the per edge values
# colour bar shows TE_ratio
pheatmap::pheatmap(
  adj_matrix_subset, 
  cluster_rows = FALSE, cluster_cols = FALSE,
  # main = "ORO type clustered by TE causal network",
  cellwidth = 15, cellheight = 15,
  fontsize = 10,
  legend_labels = TRUE,
  angle_col = "90",
  width = 10, height = 6,
  filename = here::here("figures/supplemental/TE_network_oro_cluster_heatmap2.pdf"),
)
dev.off()


```


Visualize the networks of each oro type cluster
```{r Get the dendographs corresponding to each oro type cluster}
# Cut into 5 clusters
n_clusters <- 5
oro_type_clusters <- cutree(hc, k = n_clusters)

# Create a data.frame for lookup
cluster_df <- data.frame(
  oro_type = names(oro_type_clusters),
  cluster = oro_type_clusters
)


# List to hold one meta-graph per cluster
meta_graphs <- vector("list", n_clusters)
cluster_names <- vector("list", n_clusters)

for (k in 1:n_clusters) {
  # Get oro_types in this cluster
  types_in_cluster <- cluster_df %>%
    filter(cluster == k) %>%
    pull(oro_type)
  
  # Filter TE results for these oro_types and significant edges
  cluster_te <- TE_results %>%
    filter(oro_type %in% types_in_cluster, 
           TE_pval < 0.05,
           !is.na(TE_ratio),
           0 < TE_ratio
           )
  
  # Aggregate TE_ratio per NodeX → NodeY pair
  edge_summary <- cluster_te %>%
    group_by(NodeX, NodeY) %>%
    summarise(
      n_sig = n(),
      avg_ratio = mean(TE_ratio, na.rm = TRUE),
      .groups = 'drop'
    )%>%
    mutate(
      relative_nsig = n_sig / max(n_sig)
    )
  
  # Build igraph object
  g <- graph_from_data_frame(edge_summary, directed = TRUE)
  
  # Set edge weights and labels
  E(g)$weight <- edge_summary$avg_ratio
  E(g)$width <- log1p(E(g)$weight) * 1.5
  E(g)$color <- scales::alpha("black", edge_summary$relative_nsig)
  
  meta_graphs[[k]] <- g
  cluster_names[[k]] <- paste(types_in_cluster, collapse = ", ")
}






## Rough plotting routine -- better one in next chunk

require(grid)
require(gridExtra)
require(png)

# Function to plot a graph with a title
capture_plot_as_raster <- function(g, k) {
  # Create a temporary PNG file
  tmpfile <- tempfile(fileext = ".png")
  
  # Open PNG device
  png(tmpfile, width = 800, height = 800, res = 150)
  
  # Base plot (to file)
  plot(g,
       main = paste(cluster_names[[k]]),
       edge.arrow.size = 0.4,
       # edge.label = E(g)$label,
       edge.color = "black",
       vertex.label.cex = 1,
       layout = layout_with_fr(g))
  
  dev.off()
  
  # Read and convert to rasterGrob
  img <- readPNG(tmpfile)
  file.remove(tmpfile)
  rasterGrob(img, interpolate = TRUE)
}


# Capture each plot
plots <- lapply(1:n_clusters, function(k) capture_plot_as_raster(meta_graphs[[k]], k))


# Arrange all plots
pdf(here::here("figures/supplemental/cluter_level_networks.pdf"),
    width = 10, height = 10)
gridExtra::grid.arrange(grobs = plots, ncol = 2)

dev.off()
```

```{r plot cluster level networks as ggplot plotted with dendogram}
library(ggraph)
library(tidygraph)
library(ggplot2)

nodeTextSize <- 3
titleSize <- 10

nodeColPal <- c(
  "publications"="#377eb8",
  "policy"="#984ea3",
  "legislation"="#984ea3",
  "public interest"="#ff7f00",
  "public interest weighted"="#ff7f00",
  "public support"="#ff7f00",
  "deployment"="#e41a1c"
)

# Convert network to ggplot
plot_ggraph_network <- function(g, cluster_id, cluster_label) {
  g_tidy <- as_tbl_graph(g)
  
  ggraph(g_tidy, layout = "circle") +  # You can try "kk" or "circle" or "fr" for spacing
    geom_edge_link(aes(width = 1/log1p(weight), alpha = relative_nsig),
                   arrow = arrow(length = unit(5, 'mm')),
                   color = "black", show.legend = FALSE) +
    geom_node_point(size = 6, aes(color = name)) +
    scale_color_manual(values = nodeColPal, guide="none")+
    geom_node_label(aes(label = name), repel = TRUE, #nudge_x=0.1, nudge_y = 0.1
                   size = nodeTextSize, alpha=0.8
                   ) +
    labs(title = paste(cluster_label)) +
    theme_void() +
    theme(
      plot.title = element_text(size = titleSize, hjust = 0.5),
      plot.margin = unit(c(5,5,5,5), units = "mm")
      )
}

cluster_order <- c(3,4,2,5,1)#1:n_clusters
ggraph_plots <- lapply(cluster_order, function(k) {
  plot_ggraph_network(meta_graphs[[k]], k, cluster_names[[k]])
})


library(ggdendro)

# Convert to dendrogram object
dend <- as.dendrogram(hc)
dend_data <- dendro_data(dend)

# Basic dendrogram as ggplot
dend_plot <- ggplot(segment(dend_data)) +
  geom_segment(aes(x=x, y=y, xend=xend, yend=yend)) +
  theme_void() +
  geom_text(aes(x = x, y = y, label = label, angle = -90, hjust = 0), data= label(dend_data), angle = 45, hjust=1)+
  # scale_y_continuous(expand = c(0.3, 0))+
  # scale_x_continuous(expand = c(0.1, 0))+
  coord_cartesian(clip="off")+
  labs(y="",x="")+
  theme(
    axis.text = element_blank(), 
    axis.ticks = element_blank(),
    plot.margin = unit(c(0,0,2,1), units = "cm")
    ) 
dend_plot

library(patchwork)

# Arrange the ggraph plots horizontally according to cluster number
networks_row <- wrap_plots(ggraph_plots, nrow = 1)

# Combine vertically with the dendrogram
final_plot <- dend_plot / networks_row + plot_layout(heights = c(1, 3))

# Show or save

ggsave(filename=here::here("figures/supplemental/cluster_networks_aligned.pdf"), final_plot, width = 20, height = 8)

```



```{r group oro networks conceptually and plot combined networks as ggplot plotted with dendogram}
library(ggraph)
library(tidygraph)
library(ggplot2)


## Get diagraph objects for each defined cluster
# Create a data.frame for lookup
concept_cluster_df <- data.frame(
  oro_type = c(
    "CCS","CDR-General","CDR-BC","CDR-BioPump","CDR-OAE","Efficiency","MRE-Bio","MRE-Located", "MRE-Ocean","MRE-General"
  ),
  cluster = c(1,rep(2,4),3,rep(4,4))
)

n_clusters <- length(unique(concept_cluster_df$cluster))

# List to hold one meta-graph per cluster
concept_meta_graphs <- vector("list", n_clusters)
concept_cluster_names <- vector("list", n_clusters)

for (k in 1:n_clusters) {
  # Get oro_types in this cluster
  types_in_cluster <- concept_cluster_df %>%
    filter(cluster == k) %>%
    pull(oro_type)
  
  # Filter TE results for these oro_types and significant edges
  cluster_te <- TE_results %>%
    filter(oro_type %in% types_in_cluster, 
           TE_pval < 0.05,
           !is.na(TE_ratio),
           0 < TE_ratio
           )
  
  # Aggregate TE_ratio per NodeX → NodeY pair
  edge_summary <- cluster_te %>%
    group_by(NodeX, NodeY) %>%
    summarise(
      n_sig = n(),
      avg_ratio = mean(TE_ratio, na.rm = TRUE),
      .groups = 'drop'
    )%>%
    mutate(
      relative_nsig = n_sig / max(n_sig)
    )
  
  # Build igraph object
  g <- graph_from_data_frame(edge_summary, directed = TRUE)
  
  # Set edge weights and labels
  E(g)$weight <- edge_summary$avg_ratio
  E(g)$width <- log1p(E(g)$weight) * 1.5
  E(g)$color <- scales::alpha("black", edge_summary$relative_nsig)
  
  concept_meta_graphs[[k]] <- g
  concept_cluster_names[[k]] <- paste(types_in_cluster, collapse = ", ")
}






## Plot the networks of the oro types


nodeTextSize <- 3
titleSize <- 10

nodeColPal <- c(
  "publications"="#377eb8",
  "policy"="#984ea3",
  "legislation"="#984ea3",
  "public interest"="#ff7f00",
  "public interest weighted"="#ff7f00",
  "public support"="#ff7f00",
  "deployment"="#e41a1c"
)

# Convert network to ggplot
plot_ggraph_network <- function(g, cluster_id, cluster_label) {
  g_tidy <- as_tbl_graph(g)
  
  ggraph(g_tidy, layout = "circle") +  # You can try "kk" or "circle" or "fr" for spacing
    geom_edge_link(aes(width = 1/log1p(weight), alpha = relative_nsig),
                   arrow = arrow(length = unit(5, 'mm')),
                   color = "black", show.legend = FALSE) +
    geom_node_point(size = 6, aes(color = name)) +
    scale_color_manual(values = nodeColPal, guide="none")+
    geom_node_label(aes(label = name), repel = TRUE, #nudge_x=0.1, nudge_y = 0.1
                   size = nodeTextSize, alpha=0.8
                   ) +
    labs(title = paste(cluster_label)) +
    theme_void() +
    theme(
      plot.title = element_text(size = titleSize, hjust = 0.5),
      plot.margin = unit(c(5,5,5,5), units = "mm")
      )
}

cluster_order <- 1:n_clusters
concept_ggraph_plots <- lapply(cluster_order, function(k) {
  plot_ggraph_network(concept_meta_graphs[[k]], k, concept_cluster_names[[k]])
})




library(patchwork)

# Arrange the ggraph plots horizontally according to cluster number
networks_row <- wrap_plots(concept_ggraph_plots, nrow = 2)

# Combine vertically with the dendrogram
final_plot <- dend_plot / networks_row + plot_layout(heights = c(1, 3))

# Show or save

ggsave(filename=here::here("figures/supplemental/conceptual_cluster_networks.pdf"), networks_row, width = 10, height = 8)

```



For clusters 3, 4 and 1, where there are more complicated interactions with deployment, what are the impacts of increasing public interest, publications or policy/legislation on deployment?


## Build a Probabilistic Graphical Model


🔧 Description:
Fit a Bayesian Network or Structural Equation Model based on the TE graph structure:

Use TE significance as prior knowledge.

Let the model estimate edge strengths (conditional probabilities or regression weights).

🧠 Why:
Statistically grounded way to model uncertainty and dependencies beyond pairwise effects.

Try 1 : baysean network. But I'm not convinced
```{r build graphical model - Baysean network}

library(bnlearn)
library(dplyr)
library(gRain)


# # Functions to simulate counterfactuals
# simulate_counterfactual <- function(fit) {
#   grain_bn <- as.grain(fit)
#   grain_cf <- setEvidence(grain_bn, nodes = "policy", states = "high")
#   querygrain(grain_cf, nodes = "actions", type = "distribution")$actions
# }
# 



# Format data

oroTypes <- unique(allComponentDat_model$oro_type)
results_list <- list()

for (oro in oroTypes) {
  
  
  sig_te_edges <- TE_results_summary %>%
    filter(TE_pval <= 0.05, TE_XCauseY == TRUE,!is.na(TE_ratio), oro_type == oro) %>%
    select(from=NodeX, to=NodeY) 
  sig_te_edges <- apply(sig_te_edges, 2, function(x){gsub(" ","_", x)})
  sig_te_edges <- as.data.frame(
    sig_te_edges
  )
    # distinct(NodeX, NodeY)
  
  sig_nodes <- as.character(unique(c(sig_te_edges$from, sig_te_edges$to)))
  if(length(sig_nodes)<1){
    next
  }
  
  dag_from_te <- empty.graph(nodes = gsub(" ","_", sig_nodes))
  arcs(dag_from_te) <- as.matrix(sig_te_edges)
  
  # Prepare time series
  data_oro <- allComponentDat_model %>% 
    filter(oro_type == oro) %>%
    pivot_wider(id_cols = c(year), names_from = component, values_from = y) 
  colnames(data_oro) <- gsub(" ","_", colnames(data_oro))
  data_oro <- data_oro[,sig_nodes]
  # data_oro[is.na(data_oro)] <- 0
  
  # remove obs with only NA
  data_oro <- data_oro[apply(data_oro, 1, function(y) !all(is.na(y))),]
  
  # discretize
  data_oro <- discretize(data_oro, method = "interval", breaks=10)


  bn_fit <- bn.fit(dag_from_te, data = data_oro, method = "mle")
  grain_bn <- as.grain(bn_fit)
  # # or bootstrapped
  # boot_res <- boot.strength(data_oro[,sig_nodes], R = 500, algorithm = "hc", algorithm.args = list(whitelist = arcs(dag_from_te)))
  # bootstrap_fits <- averaged.network(boot_res, threshold = 0.85)  #only keep strong edges
  # # simulate counterfactual
  # fits <- lapply(bootstrap_fits$nodes, function(dag) bn.fit(dag, data_oro))
  # cf_dists <- lapply(fits, simulate_counterfactual)
  # grain_bn <- as.grain(fits[[1]])
  
  # simulate cf
  # Set evidence (intervene on policy)
  grain_cf <- setEvidence(grain_bn, nodes = "policy", states = "high")

  # Get updated beliefs for deployment
  cf_deployment <- querygrain(grain_cf, nodes = "deployment", type = "marginal")

  # Baseline belief (no intervention)
  baseline_deployment <- querygrain(grain_bn, nodes = "deployment", type = "marginal")
  
  # Compate baseline vs cf
  cf <- cf_deployment$deployment
  base <- baseline_deployment$deployment
  print(base-cf)
  
  # 7. Loop over interventions on each node → observe effect on "deployment"
  # Since you’re simulating an intervention (e.g. P(deployment | do(policy = high))), what you want is the marginal distribution of deployment given the intervention.
  all_nodes <- nodes[nodes != "deployment"]
  cf_results <- list()

  for (node in all_nodes) {
    for (state in c("high")) { # c("low","medium","high")
      # Set evidence and query deployment
      bn_cf <- setEvidence(grain_bn, nodes = node, states = state)
      cf_deploy <- querygrain(bn_cf, nodes = "deployment", type = "marginal")
      cf_results[[paste(node, state, sep = "_")]] <- cf_deploy
    }
  }
  
  # baseline (no intervention)
  baseline <- querygrain(grain_bn, nodes = "deployment", type = "marginal")
  
  # Compile
  results_list[[oro]] <- list(
    nodes = nodes,
    dag = dag,
    fit = fit,
    baseline = baseline,
    counterfactuals = cf_results
  )

 
}




## Visualize
cfDeltas <- data.frame()
for(oro in oroTypes){
  cf <- results_list[[oro]]$counterfactuals
  baseline <- results_list[[oro]]$baseline
  
  if(is.null(baseline)){
    next
  }
  
  # Step 1: Get bin names and extract midpoints
  sample_probs <- cf_df$Probability[1]$deployment
  bin_names <- names(sample_probs)
  
  # Extract numeric midpoints from bin labels like "(4.7,8.4]"
  get_midpoint <- function(bin_label) {
    nums <- as.numeric(str_extract_all(bin_label, "[0-9.]+")[[1]])
    mean(nums)
  }
  
  bin_midpoints <- sapply(bin_names, get_midpoint)
  
  # Step 2: Get baseline EV
  baseline_probs <- baseline$deployment[bin_names]
  baseline_ev <- sum(baseline_probs * bin_midpoints)
  
  # Step 3: Compute expected values from counterfactuals
  cf_ev_df <- map_dfr(seq_along(cf_df$Probability), function(i) {
    p <- cf_df$Probability[i]$deployment[bin_names]
    ev <- sum(p * bin_midpoints)
    tibble(
      Intervention = cf_df$Intervention[i],
      ExpectedValue = ev
    )
  }) %>%
    mutate(DeltaEV = ExpectedValue - baseline_ev)
  cf_ev_df$oro_type = oro
  
  cfDeltas <- cfDeltas %>% bind_rows(cf_ev_df)
}

summary(cfDeltas) # does'nt work, all the deltas are 0

# ggplot(cf_ev_df, aes(x = reorder(Intervention, DeltaEV), y = DeltaEV, fill = DeltaEV > 0)) +
#   geom_col(show.legend = FALSE) +
#   coord_flip() +
#   labs(
#     title = "Change in Expected Deployment from Interventions",
#     x = "Intervention",
#     y = expression(Delta*" Expected Deployment")
#   )


```

```{r build qualitative network model for each oro}
# remotes::install_github("SWotherspoon/QPress")
library(QPress)
library(reshape2)
nSims <- 5000

source(here::here("R/impact.barplot.myMod"))
source(here::here("R/extend.vector"))

addTaskCallback(function(...) {set.seed(123);TRUE})

# Format data

oroTypes <- unique(allComponentDat_model$oro_type)
qp_oro_results_df <- data.frame()


# maybe instead of by oro type, make edge weight proportional to agreement/consensus...
# only simulate if deployment is present
for (oro in oroTypes) {
  
  sig_te_edges <- TE_results %>%
    filter(TE_pval <= 0.05, TE_XCauseY == TRUE,!is.na(TE_ratio), oro_type == oro) %>%
    select(From=NodeX, To=NodeY) %>%
    mutate(
      Type = "P"
    )
  
  
  sig_nodes <- as.character(unique(c(sig_te_edges$From, sig_te_edges$To)))
  if(length(sig_nodes)<1){
    next
  }
  if(!("deployment" %in% sig_nodes)){
    next
  }
  
  sig_te_edges <- sig_te_edges %>%
    mutate(
      From = factor(From, levels = sig_nodes),
      To = factor(To, levels=sig_nodes),
      Type = factor(Type, levels = c("N","P","U","Z")),
      Group = 0
    )
  sig_te_edges$Pair <- 1:nrow(sig_te_edges)
  
  allEdgesLabels <- edge.labels(sig_te_edges) 
  
  Mod <- retain.groups(sig_te_edges, groups=0) # all certain interactions
  Mod <- enforce.limitation(sig_te_edges) # enforce self-limiting effect
  ModSim <- system.simulate(nSims, Mod) # simulate
  
  
  
  # Loop over interventions on each node → observe effect on "deployment"
  all_nodes <- sig_nodes[sig_nodes != "deployment"]
  # temp <- impact.table(ModSim)

  for (node in all_nodes) {
    press <- c(1)
    names(press) <- node

    # Set evidence and query deployment
    temp <- impact.barplot.myMod(ModSim,
        perturb = press,
        plot=FALSE, percentage = TRUE
    )
    temp <- as.data.frame(t(temp["deployment",]))
    temp$oro_type <- oro
    temp$Press_perturb <- node
    
    qp_oro_results_df <- qp_oro_results_df %>% bind_rows(temp)
    
  }

 
}


qp_oro_results_df %>% 
  group_by(Press_perturb) %>%
  summarise(
    avg_pos = mean(Positive, na.rm=T)
  )
#   Press_perturb            avg_pos
#   <chr>                      <dbl>
# 1 legislation                  0  
# 2 policy                      25  
# 3 public interest             66.7
# 4 public interest weighted    33.3
# 5 public support              33.3
# 6 publications                75 


# Plot results
# For each ORO type

ggplot(qp_oro_results_df, 
       aes(x=Press_perturb, y=oro_type, fill = Positive))+
  geom_tile()+
  scale_fill_steps(name = stringr::str_wrap("Positive effect on deployment (% sims)", width=25), breaks = c(50,100), labels = c("0","100"))+
  labs(y="ORO type",x="Press perturbation")+
  theme_bw()+
  theme(
    legend.position = "bottom",
    axis.text.x=element_text(angle=45, hjust=1)
  )


ggsave(
  here::here("figures/supplemental/qp_oro-level_heatmap.pdf"),
  width=5, height=4, units="in"
)
```


But from the conceptual cluster networks, only CDR and MRE groups have deployment in the network. SO look at these clusters
```{r build qualitative network model for CDR and MRE conceptual clusters}
# remotes::install_github("SWotherspoon/QPress")
library(QPress)
library(reshape2)
nSims <- 5000

source(here::here("R/impact.barplot.myMod"))
source(here::here("R/extend.vector"))
source(here::here("R/my.QPress.R"))

addTaskCallback(function(...) {set.seed(123);TRUE})

# Format data
weightSimulation = FALSE
uncertainThreshold = 1 # the threshold of agreement below which the edge is sample with a probability weight of 0.5
oroTypes <- unique(allComponentDat_model$oro_type)
qp_concept_cluster_results_df <- data.frame()

# Create a data.frame for lookup
concept_cluster_df <- data.frame(
  oro_type = c(
    "CCS","CDR-General","CDR-BC","CDR-BioPump","CDR-OAE","Efficiency","MRE-Bio","MRE-Located", "MRE-Ocean","MRE-General"
  ),
  cluster = c(1,rep(2,4),3,rep(4,4))
)

n_clusters <- length(unique(concept_cluster_df$cluster))

# List to hold one meta-graph per cluster
concept_meta_graphs <- vector("list", n_clusters)
concept_cluster_names <- vector("list", n_clusters)




# maybe instead of by oro type, make edge weight proportional to agreement/consensus...
# only simulate if deployment is present
for (k in 1:n_clusters) {
  # Get oro_types in this cluster
  types_in_cluster <- concept_cluster_df %>%
    filter(cluster == k) %>%
    pull(oro_type)
  
  
  
  sig_te_edges <- TE_results %>%
    filter(TE_pval <= 0.05, TE_XCauseY == TRUE,!is.na(TE_ratio), oro_type %in% types_in_cluster) %>%
    group_by(NodeX, NodeY) %>%
    summarise(
      n_sig = n(),
      .groups = 'drop'
    )%>%
    mutate(
      Probability = n_sig / max(n_sig),
      Group = ifelse(uncertainThreshold <= Probability,0,1)
    ) %>%
    arrange(desc(Probability)) %>%
    select(From=NodeX, To=NodeY, Group, Probability) %>%
    mutate(
      Type = "P"
    )
  
  
  sig_nodes <- as.character(unique(c(sig_te_edges$From, sig_te_edges$To)))
  if(length(sig_nodes)<1){
    next
  }
  if(!("deployment" %in% sig_nodes)){
    next
  }
  
  sig_te_edges <- sig_te_edges %>%
    mutate(
      From = factor(From, levels = sig_nodes),
      To = factor(To, levels=sig_nodes),
      Type = factor(Type, levels = c("N","P","U","Z")),
      # Group = factor(Group)
    )
  sig_te_edges$Pair <- 1:nrow(sig_te_edges)
  
  allEdgesLabels <- edge.labels(sig_te_edges) 
  
  # Mod <- retain.groups(sig_te_edges, groups=0) # all certain interactions
  Mod <- enforce.limitation(sig_te_edges%>% select(-Probability)) # enforce self-limiting effect
  # s<- community.sampler(Mod)
  # s$select(0.5)
  # ModSim <- system.simulate(100, Mod,sampler = s) # simulate
  # print(impact.table(ModSim))
  
  if(weightSimulation){
    uncertainSampleProbsDf <- sig_te_edges %>%
    filter(Group == 1) %>%
    select(Pair, Probability) %>%
    distinct(Pair,.keep_all = TRUE)
  
    ModSim <- my.system.simulate(nSims, Mod,required.groups = c(0), uncertainSampleProbsDf=uncertainSampleProbsDf)
  }else{
    ModSim <- system.simulate(100, Mod)
  }
  
  # Loop over interventions on each node → observe effect on "deployment"
  all_nodes <- sig_nodes[sig_nodes != "deployment"]
  # temp <- impact.table(ModSim)

  for (node in all_nodes) {
    press <- c(1)
    names(press) <- node

    # Set evidence and query deployment
    temp <- impact.barplot.myMod(ModSim,
        perturb = press,
        plot=FALSE, percentage = TRUE
    )
    temp <- as.data.frame(t(temp["deployment",]))
    temp$cluster <- paste(types_in_cluster, collapse = ", ")
    temp$Press_perturb <- node
    
    qp_concept_cluster_results_df <- qp_concept_cluster_results_df %>% bind_rows(temp)
    
  }

 
}


qp_concept_cluster_results_df %>% 
  group_by(Press_perturb) %>%
  summarise(
    avg_pos = mean(Positive, na.rm=T)
  )



# Plot results
# For each ORO type

ggplot(qp_concept_cluster_results_df %>% filter(Positive>0), 
       aes(x=Press_perturb, y=stringr::str_wrap(cluster, width=10), fill = Positive))+
  geom_tile()+
  scale_fill_stepsn(name = stringr::str_wrap("Positive effect on deployment (% sims)", width=25), n.breaks=5, colours = viridis(5))+
  labs(y="ORO type cluster",x="Press perturbation")+
  theme_bw()+
  theme(
    legend.position = "bottom",
    axis.text.x=element_text(angle=45, hjust=1)
  )


ggsave(
  here::here("figures/supplemental/qp_concept_cluster_heatmap.pdf"),
  width=5, height=4, units="in"
)
```





# Changepoint sub-analysis

For efficiency and CCS, Efficiency and MRE-Ocean, looks like change points in deployment time series -- what could have caused these increases? Could they be related to change points in other time series? While we cannot infer causality, we can say that these change points occur around the same time...

To explore this, fit a Bayesian multiple change point model (adapted from Cahill et al. 2015, doi: 10.1088/1748-9326/10/8/084002 for different distributions depending on response variable: ZIP for legislation and policy documents, Poisson for deployments for CDR-OAE, number of publications and number of posts, and Gamma for other deployment metrics and number of weighted posts as they are non-negative continuous data). This model estimates the number of probable change points, and then for each change point the posterior distribution. 

```{r changepoint analysis, eval=FALSE}

# The oro types to analyse 
oroTypes = c("CCS","Efficiency","MRE-Ocean","CDR-OAE")
# The different time series components to analyse (e.g. deployment, legislation, policy, posts, etc)
components <- unique(allComponentDat_model$component)

# Jags initialization parameters
jagsInits <- list(
  # Maximum number of changepoints to look for
  "K_max"=2,
  # Time buffer for looking for change points. 
  # i.e. don't find a change point at the first or last year
  "cpmin"="jagsData$MINX+1", 
  "cpmax"="jagsData$MAXX-1")

# Compile all data into a list to iterate through
dataList <- list()
for(c in components){
  dataList[[c]] <- allComponentDat_model %>% filter(oro_type %in% oroTypes, component == c)
 
}

# Compile all model inputs into a list to iterate through
modelInputs <- list(
  "data" = dataList,
  "bugsModFiles" = list(
    "deployment" = "R/bugs-models/GammaModMultCP",
    "publications" = "R/bugs-models/PoissonModMultCP",
    "policy" = "R/bugs-models/ZIPModMultCP",
    "legislation" = "R/bugs-models/ZIPModMultCP",
    "public interest" = "R/bugs-models/PoissonModMultCP",
    "public interest weighted" = "R/bugs-models/GammaModMultCP",
    "public support" = "R/bugs-models/GammaModMultCP"
  ),
  "jagsData" = list(jagsInits)[rep(1,length(components))]
)



## Loop through all oro types and time series (components) to run change point analysis
cutpoint_results <- data.frame()
cutpoint_densities <- data.frame()

for(oro in oroTypes){
  # oro = "CCS"
  
  for(i in 1:length(components)){
    # i = 4
      
    print(paste(oro,":", components[i]))
    
    # get model to fit
    modFile <- modelInputs$bugsModFiles[[components[i]]]
    
    # If OAE deployment, y is poisson not Gamma so change
    if(components[i] == "deployment" & oro %in% c("CDR-OAE")){
      modFile <- "R/bugs-models/PoissonModMultCP"
    }
    
    # Format data
    jagsData <- modelInputs$jagsData[[i]]
    myinits = list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 123)
    dat <- modelInputs$data[[components[i]]][modelInputs$data[[components[i]]]$oro_type == oro,]
    if(nrow(dat) < 5){
      print("Insufficient data")
      next
    }
    yearLims <- range(modelInputs$data[[i]][
      modelInputs$data[[components[i]]]$oro_type == oro,"year"
    ], na.rm=T)
    
    # Format specific inputs depending on the model
    if(grepl("pubBinom", modFile)){ 
      trialVar <- ifelse(grepl("ORO", modelInputs$y_variable[[i]]),
                         "n_OC","total_posts")
      successVar <- ifelse(grepl("ORO", modelInputs$y_variable[[i]]),
                         "n_ORO","n_posts")
      # if a proportion, only keep proportions of total values above 100 # modelInputs$y_variable[[i]] == "prop_ORO"
      keepYears <- dat$year[dat$variable_name == trialVar & 100 < dat$y]
      dat <- dat[dat$year %in% keepYears,]
      dat <- dat %>% arrange(variable_name, year)
      nTrial <- dat$y[dat$variable_name == trialVar]
      jagsData$nTrial <- nTrial
      jagsData$nSuccess <- dat$y[dat$variable_name == successVar]
      # datCast <- reshape2::dcast(dat, ... ~ variable_name, value.var = "y")
    }
    
    if(grepl("ZIP", modFile)){ 
      # if the model is zero inflated poisson, 
      # Fill missing years with 0 values
      dat <- dat %>% tidyr::complete(year = yearLims[1]:yearLims[2])
      dat$y[is.na(dat$y)] <- 0
      dat <- dat %>% tidyr::fill(component, oro_type, variable_name)
      # add the latent binomial indicator, as to whether there are no documents at all
      myinits$w <- dat$y
      myinits$w[myinits$w > 0] <- 1
      
    }
    
    if(grepl("Poisson", modFile)){ 
      # Fill missing years with 0 values
      dat <- dat %>% tidyr::complete(year = yearLims[1]:yearLims[2])
      dat$y[is.na(dat$y)] <- 0
      dat$y <- round(dat$y) # force discrete y values
      dat <- dat %>% tidyr::fill(component, oro_type, variable_name)
      
    }
    
    if(grepl("Gamma", modFile)){
      # Fill missing years with 0 values
      dat <- dat %>% tidyr::complete(year = yearLims[1]:yearLims[2])
      dat$y[is.na(dat$y)] <- 0
      dat$y[dat$y==0] <- 0.1
      dat <- dat %>% tidyr::fill(component, oro_type, variable_name)
    }
    
    # make sure no NAs
    dat <- na.omit(dat)
    
    # Skip if after subsetting data there is insufficient data points
    if(nrow(dat) < 5){
      print("Insufficient data after subsetting")
      next
    }
    #with(dat, plot(year, y))
    
    jagsData$y <- dat$y
    jagsData$x <- dat$year 
    jagsData$MINX <- min(jagsData$x)
    jagsData$MAXX <- max(jagsData$x)
    jagsData$cpmin <- eval(parse(text = jagsData$cpmin))
    jagsData$cpmax <- eval(parse(text = jagsData$cpmax))
    
    if((jagsData$cpmax-jagsData$cpmin)<5){
      jagsData$cpmin <- jagsData$MINX
      jagsData$cpmax <- jagsData$MAXX
    }
    
    # jagsData$log_y <- log(dat$y)
    # jagsData$N <- length(dat$y)
    
    ## Fit mdel
    modFile <- "R/bugs-models/GammaModMultCP"
    bugs.model <- readChar(modFile, file.info(modFile)$size)
    
    jagsModel <- tryCatch(
      {
        rjags::jags.model(
          file = textConnection(bugs.model),
          data = jagsData,
          inits = myinits,
          n.chains = 5,
          n.adapt = 1500,
          quiet = FALSE
        )
        },
     error = function(e){
       return(NULL)
     }
    )
    if(is.null(jagsModel)){
      next
    }
    
    
    update(jagsModel, 1000) # burnin
    s = rjags::coda.samples(
          model = jagsModel,
          variable.names = c("alpha","beta","K","x_cp","z","pk"),
          n.iter = 1000,
          quiet = FALSE
        )
    
    qs <- summary(s)$quantile
    # qs
    # plot(s[,"pk[2]"]) # check mixing
    
    # Subset to only the change points that are likely 
    # And if there are likely change points, only those that indicate a positive change
    K <- round(quantile(qs["K",], 0.5))
    if(grepl("ZIP", modFile)){
      K <- sum(0.5 <= qs[grepl("pk", rownames(qs)),"50%"])
    }
    
    if(K == 0){
      print("No significant change points")
      next
    }else{
      pk <- qs[grep("pk", rownames(qs)),"50%"] # probability for each change point
      x_cp <- qs[grep("x_cp", rownames(qs)),"50%"] # locations of each change point
      # arrange in order of decreasing probability to find the most K probable
      pkIndKeep <- order(pk, decreasing=T) 
      pkIndKeep <- pkIndKeep[1:K] 
      # then re-arrange in chronological order for calculating trends of segments
      pkIndKeep <- sort(pkIndKeep) 
      
      # Find which change points mark a positive change in trend
      trends <- vector("numeric", K+1) # store the trend for each segment
      
      # starts <- c(min(jagsData$x), round(x_cp[pkIndKeep])) 
      starts <- c(min(jagsData$x), sort(round(x_cp[pkIndKeep])))
      match(starts, round(x_cp[pkIndKeep]))
      ends <- c(round(x_cp[pkIndKeep]), max(jagsData$x))
      for(k in 1:(K+1)){
        if(grepl("Binom", modFile)){
          trends[k] <- jagsData$y[which.min(abs(jagsData$x-ends[k]))] - jagsData$y[which.min(abs(jagsData$x-starts[k]))]
        }else if(grepl("ZIP|Pois|Gamma", modFile)){
          trends[k] <- sum(jagsData$y[which.min(abs(jagsData$x-starts[k])):which.min(abs(jagsData$x-ends[k]))] )
        }
        
      }
      # only keep the positive changes in trends
      matchInd <- match(starts[-1], round(x_cp[pkIndKeep]))
      pkIndKeep <- pkIndKeep[matchInd[which(0 < diff(trends))]] 
      
      if(length(pkIndKeep) == 0){
        print("no changepoints marking increasing trend")
        next
      }else{
        pkIndNames <- names(x_cp[pkIndKeep])
        
        # # If there are two cut points within 3 years of each other, 
        # # keep most probable one 
        # if(1 < length(pkIndKeep)){
        #   if(diff(qs[pkIndNames,"50%"]) < 3){
        #     pkIndNames <- pkIndNames[1]
        #   }
        # }
        
        # save probability density of the cut points for plotting
        calc_density <- function(vals){
          dens <- density(vals, n=100)
          return(data.frame("year"=dens$x, "cp_density"=dens$y))
        }
        densTemp <- do.call(rbind, s)
        densTemp <- as.matrix(densTemp[,pkIndNames], ncol=length(pkIndNames))
        densTemp <- do.call(rbind.data.frame, apply(densTemp, 2, function(x) calc_density(x)))
        densTemp$cp_id <- rep(pkIndNames, 100)[sort(rep(1:length(pkIndNames),100))]
        
        
        # Save summary of quantiles
        dfTemp <- do.call(rbind, apply(qs[pkIndNames,,drop=FALSE], 1, function(x) as.data.frame(x)))
        dfTemp$quantile = rep(colnames(qs), length(pkIndNames))
        dfTemp$cp_id <- rep(pkIndNames, 5)[sort(rep(1:length(pkIndNames),5))] #rep(1:length(pkIndNames), 5) %>% sort
        rownames(dfTemp) <- NULL
        # Cap the distribution of cutpoints at the hard limits of the data
        colnames(dfTemp)[which(colnames(dfTemp) == "x")] <- "year"
        dfTemp$year <- ifelse(dfTemp$year < yearLims[1], yearLims[1], dfTemp$year)
        dfTemp$year <- ifelse(yearLims[2] < dfTemp$year, yearLims[2], dfTemp$year)
        # Add id variables
        addIdVariables <- function(df){
          df %>% mutate(
                        oro_type = oro,
                        component = dat$component[1],
                        variable_name = dat$variable_name[1])
        }
        dfTemp <- addIdVariables(dfTemp)
        densTemp <- addIdVariables(densTemp)
        
        ## Bind data to results
        cutpoint_results <- rbind(cutpoint_results, dfTemp)
        cutpoint_densities <- rbind(cutpoint_densities, densTemp)
        # remove jags model
        rm(jagsModel)
      }
    }

      

  } # end looping through ORO types
} # end looping through bugs models



## Save results
# save(cutpoint_results,cutpoint_densities, file=here::here("outputs/cutpointResults_mitigation.RData"))
```


CP attribution - if a deployment cutpoint is identified what other cutpoints happened within -4 years + 2 years?
```{r CP attribution}
# Load data from previous chunk
load(here::here("outputs/cutpointResults_mitigation.RData"))

oroTypes <- c("CCS","Efficiency","MRE-Ocean","CDR-OAE") 

attrCP_summary <- data.frame()
for(oro in oroTypes){
  depCP <- cutpoint_results %>%
    filter(component == "deployment", oro_type == oro, quantile == "50%")
  
  # For each deployment cutpoint, which other cutpoints ocurred within pm 2 years?
  for(cpid in unique(depCP$cp_id)){
    targetYear <- round(depCP$year[depCP$cp_id == cpid])
    attrCPs <- cutpoint_results %>%
      filter(component != "deployment", oro_type == oro, quantile == "50%",
             # cp_id == cpid,
             round(year) <= targetYear+3 & targetYear-3 <= round(year))
    
    attrCP_summary <- attrCP_summary %>%
      bind_rows(attrCPs)
    
  }
  
}

# View(cutpoint_results)
print(attrCP_summary)



```


When big changes happen, they co-occur with change in legislation and public interest/support...

```{r plot of cutpoints}

yBackTransform <- function(value, ogRange, newRange){
  b <- diff(newRange) / diff(ogRange)
  a <- newRange[1] - b * ogRange[1]
  return(a + b * value)
}
  

# Get raw data to plot
cpDeployPlotData <- data.frame()
cpDensityPlotData <- data.frame()
yAxisBreaks <- data.frame()
for(oro in unique(attrCP_summary$oro_type)){
  
  rawDat <- allComponentDat_model %>% 
    filter(oro_type == oro, component == "deployment") %>%
    mutate(y_scale = scales::rescale(y, c(0,1), na.rm=T)) %>%
    mutate(
      label = paste(
        scales::number(y, accuracy =1, scale_cut = append(scales::cut_long_scale(), 1, 1), 
                       big.mark = ".", decimal.mark = ",")
      )
    )
  ogRange <- range(rawDat$y, na.rm=T)
  newRange <- c(0,1)
  ogBreaks <- pretty(ogRange, n = 4)
  newBreaks <- yBackTransform(ogBreaks, ogRange, newRange)
  if(1 < ogRange[2]){
    ybl <- scales::number(ogBreaks, accuracy =1, scale_cut = append(scales::cut_long_scale(), 1, 1), 
                       big.mark = ",", decimal.mark = ".")
  }else{
    ybl <- scales::number(ogBreaks, accuracy =0.001, scale_cut = append(scales::cut_short_scale(), 1, 1), 
                       big.mark = ",", decimal.mark = ".")
  }
  
  yAxisBreaks <- yAxisBreaks %>%
    bind_rows(
      data.frame(
        y_breaks = ogBreaks,
        y_breaks_label = ybl,
        y_scale_breaks = newBreaks,
        oro_type = oro,
        component = "deployment"
      )
    )
  
  cpDens <- cutpoint_densities %>%
    filter(component == "deployment") %>%
    mutate(cp_group = paste(oro, component, cp_id)) %>%
    group_by(cp_group) %>%
    mutate(
      cp_density = scales::rescale(cp_density, to=c(0,1))
    )
  
  cpDeployPlotData <- cpDeployPlotData %>% bind_rows(rawDat)
  cpDensityPlotData <- cpDensityPlotData %>% bind_rows(cpDens)
 
}

stripLabsDF <- cpDeployPlotData %>%
  distinct(oro_type, variable_name) %>%
  mutate(
    labels = paste(oro_type, variable_name, sep="\n")
  )
stripLabs <- stripLabsDF$labels
names(stripLabs) <-stripLabsDF$oro_type


changePoint_ggp <- ggplot()+
  geom_ribbon(data = cpDensityPlotData, 
       aes(x=.data$year, ymax=.data$cp_density, ymin=0, group=.data$cp_group), 
       fill = "lightgrey", alpha = 0.5)+ # , ymin=-Inf, 
  geom_text(data = yAxisBreaks,
            aes(x=-Inf, y=.data$y_scale_breaks, label = .data$y_breaks_label),
            size = 3, col = "black", hjust = 1, vjust=0)+
  geom_vline(data = attrCP_summary, 
             aes(xintercept=.data$year, col = .data$component, group = .data$cp_id), 
             linewidth = 1.5)+
  # geom_rect(data = cutpoint_results %>% 
  #            filter(paste(cp_id, oro_type, component) %in% with(attrCP_summary, paste(cp_id, oro_type, component))) %>%
  #            reshape2::dcast(...~quantile, value.var = "year"),
  #         aes(xmin = `25%`, xmax=`75%`, group=cp_id, fill = component), ymin=-Inf, ymax=Inf, alpha=0.5,
  #         inherit.aes = FALSE)+
  geom_line(data = cpDeployPlotData, 
            aes(x=.data$year, y=.data$y_scale),
            col = "black", linewidth = 1)+
  scale_y_continuous(
    breaks = NULL, #pretty_breaks(n = 4),
    name = ""
  )+
  scale_x_continuous(name = "Year")+
  
  # Color scales
 scale_color_manual(name="Median\nchange-point",
                  breaks = c("legislation","public interest","public interest weighted","public support","publications"),
                  labels =c("Legislation","Public interest\n(N Posts)","Public interest (w)\n(N posts weighted)","Public support\n(N positive posts weighted)","Publications"),
                  values = c("#7570b3","#1b9e77","#d95f02","#e78ac3","#377eb8")
                  )+
  # scale_fill_manual(name="Time series dataset",
  #                 breaks = c("legislation","public interest","public interest weighted","public support" ),
  #                 labels =c("Legislation","N Posts","N posts (weighted)","Public support"),
  #                 values = c("#7570b3","#1b9e77","#d95f02","#e78ac3")
  #                 )+
  guides(color = guide_legend(nrow = 2))+

  facet_wrap(vars(oro_type), ncol=1, labeller = labeller(oro_type = stripLabs))+
  coord_cartesian(clip = "off")+
  theme_minimal()+
  theme(
    axis.text.y = element_blank(),
    plot.margin = unit(c(0.5,1,1,1.5), units = "cm"),
    legend.position = "bottom"
  )

changePoint_ggp

ggsave(
  here::here("figures/main/changePointTimeSeries.pdf"),plot=changePoint_ggp,
  width = 5, height=7, units='in'
)
```





