---
title: "0.2_explore-external-data"
author: "Devi Veytia"
date: "2024-04-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
library(dplyr)
library(ggplot2)
```

# Marine energy

## Deployment scale

```{r load from databases and deduplicate}
# df1 <- readxl::read_excel(here::here("data/raw-data/external/PRIMRE-marine-energy-projects-database.xlsx"),sheet = "Projects")
# df2 <- readxl::read_excel(here::here("data/raw-data/external/Tethys-pnnl-marine-energy-metadata.xlsx"),sheet = "Projects")
# df <- rbind(df1, df2)
# 
# # Deduplicate 
# dups <- revtools::find_duplicates(data=df, match_variable = "project_name", to_lower = TRUE)
# marine_energy_df <- revtools::extract_unique_references(df, dups)
# 
# # save
# save(marine_energy_df, file = here::here("data/raw-data/external/PRIMRE-Tethys-combined.RData"))

load(here::here("data/raw-data/external/PRIMRE-Tethys-combined.RData"))

# Quick plot 
marine_energy_df <- marine_energy_df %>%
  mutate(year = as.numeric(ifelse(is.na(operational_start_year), project_start_year, operational_start_year)),
         max_rated_power_capacity_MW = as.numeric(max_rated_power_capacity_MW)) %>%
  filter(!is.na(year))
marine_energy_df$year <- as.Date(paste0(marine_energy_df$year,"-01-01"))
marine_energy_df <- marine_energy_df %>%
  tidyr::replace_na(list(max_rated_power_capacity_MW = 0)) %>%  # Replace NA with zeros
  group_by(project_life_cycle, year) %>%
  arrange(project_life_cycle, year) %>%  # Arrange by year so adding works
  group_by(project_life_cycle) %>%
  mutate(cumMW = cumsum(max_rated_power_capacity_MW))

summary(marine_energy_df)
  

ggplot(marine_energy_df, aes(year, cumMW, fill = project_life_cycle, col = project_life_cycle))+
  geom_area(alpha = 0.5)+
  scale_x_date()+
  theme_bw()
```

```{r map countries of MRE locations? -- need to do}

```

## IRENA renewable electricity statistics

```{r}
df <- readxl::read_excel(here::here("data/raw-data/external/IRENA-electricity-statistics-by-country-year.xlsx"),
                         sheet = "ELECSTAT-C", col_names = FALSE, skip = 2)
colnames(df) <- c("country","technology","data_type","grid_connection","year","value")

df <- df %>% 
  tidyr::fill(country, technology, data_type, data_type, grid_connection, .direction = "down")

# filter data to one variable, all grid connections
df <- df %>%
  filter(data_type == "Electricity Installed Capacity (MW)", grid_connection == "All", technology %in% c("Marine energy","Offshore wind energy"))

# Format columns
df$year <- as.Date(paste0(df$year,"-01-01"))

# Quick plot
ggplot(df, aes(year, value, fill = country, col = country))+
  facet_grid(technology ~., scales = 'free_y') +
  geom_area(alpha=0.5)+
  labs(y = "Electricity Installed Capacity (MW)")+
  scale_x_date()+
  theme_bw()+
  theme(legend.position = 'none')

# can colour this by country type? 

```



# Novel CDR deployments - State of CDR Report 2022

```{r load in novel CDR deployment data}
cdr_deployment <- readxl::read_excel(here::here("data/raw-data/external/State-of-Carbon-Report-2022-Novel-CDR-Deployment-Data.xlsx"), skip = 1)

carbon_sinks <- unique(do.call(c, unique(as.vector(cdr_deployment[,3:6]))))
carbon_sinks

cdr_deployment$carbon_sink <- apply(cdr_deployment[,3:6], 1, paste, collapse=", ")
cdr_deployment <- cdr_deployment %>%
  mutate(ocean = ifelse(
    grepl("Aquatic", carbon_sink) | grepl("Deep Ocean", carbon_sink)| 
      grepl("Dissolved CO2", carbon_sink) | grepl("Bicarbonate", carbon_sink), TRUE,FALSE
  ))
cdr_deployment$year = as.Date(paste0(cdr_deployment$`Year active`,"-01-01")) 

summary(cdr_deployment)

cdr_deployment %>%
  filter(ocean) %>%
  group_by(year) %>%
  summarise(`Gross Removals\n(tCO2/year)` = sum(`Gross Removals\n(tCO2/year)`, na.rm=T)) %>%
  mutate(`Gross Removals\n(tCO2/year)` = cumsum(`Gross Removals\n(tCO2/year)`)) %>%
  ggplot(aes(year, `Gross Removals\n(tCO2/year)`))+
  geom_line()+
  geom_point()+
  scale_x_date()+
  labs(y = "Cumulative Gross Removals\n(tCO2/year)",
       caption = "Source: State of CDR Report 2022 - Novel CDR Deployment Data\nfiltered for where carbon is stored in ocean sinks")+
  theme_bw()
```

```{r Ocean Visions mCDR Field Trials}
mCDR_field <- readxl::read_excel(here::here("data/raw-data/external/Ocean-visions-mCDR-field-trial-database.xlsx"),
                                 sheet = "Data")
mCDR_field_summary <- mCDR_field %>%
  filter(!is.na(`Start of Pilot`)) %>%
  group_by(`Start of Pilot`) %>%
  summarise(
    n_projects = n(),
    planned_seq = sum(as.numeric(`Planned Sequestration (tons CO2)`), na.rm = T)
  ) %>%
  mutate(year = as.Date(as.numeric(`Start of Pilot`), origin = "1899-12-30")) %>%
  arrange(year) %>%
  mutate(cumulative_projects = cumsum(n_projects),
         cumulative_seq = cumsum(planned_seq))
  

mCDR_field_summary <- reshape2::melt(mCDR_field_summary, measure.vars = c("cumulative_projects", "cumulative_seq"),
                                     id.vars = "year")

mCDR_field_summary$variable <- factor(mCDR_field_summary$variable,
                                      levels = c("cumulative_projects", "cumulative_seq"),
                                      labels = c("Projects (N)","Planned sequestration (tCO2)"))
summary(mCDR_field_summary)

ggplot(mCDR_field_summary, aes(year, value))+
  facet_grid(variable ~., scales = "free_y", labeller = label_wrap_gen(width = 25))+
  geom_line()+
  labs(y = "Cumulative value", caption = "Source: Ocean Visions mCDR field trials database")+
  scale_x_date(limits = c(as.Date("2018-01-01"), max(mCDR_field_summary$year)))+
  theme_bw()
```


# Increasing efficiency

## Fishing

Estimates of fishing effort were sourced from a global database based on estimates of total vessel engine size and number of fishing days in a year, assembled from FAO, the European Union, regional tuna-management bodies and other sources50,51

Fuel consumption rates were extracted from FEUD19 (litres per round weight tonne). Records of fisheries operating before 1985 were excluded from analysis, as were any records for which target species group or gear type could not be determined

The combined rate of emissions was 3.3 kg CO2-eq GHG per litre of fuel combusted.


```{r fuel use intentisty from greer et al 2019}

# Emissions intensity
dat <- metaDigitise::metaDigitise(dir=here::here("data/raw-data/external"), summary = FALSE)

greer_industrial <- dat$scatterplot$`Greer-et-al-2019_emissions-intensity-figure.jpg`
greer_industrial <- greer_industrial %>%
  filter(1950 <= x & 0 <= y)

save(greer_industrial, file= here::here("data/raw-data/external/Greer-et-al-2019_emissions-intensity_industrial.RData"))


# Fuel coefficient (engine efficiency)
greer_FC <- readxl::read_excel(here::here("data/raw-data/external/Greer-et-al-2019_fuel-coefficient-table-5.xlsx"), sheet = "fuel_coefficient")
greer_FC <- greer_FC %>%
  mutate(variable = "Fuel coefficient") %>%
  rename(y=fuel_coefficient)%>%
  select(variable, year, y)
  


# Join dataframes
fishing_efficiency_dat <- greer_industrial %>%
  mutate(year = round(x)) %>%
  mutate(variable = "Emissions intensity (tCO2/tcatch)") %>%
  select(variable, year, y) 
  
fishing_efficiency_dat <- rbind(fishing_efficiency_dat,greer_FC)
fishing_efficiency_dat$year <- as.Date(paste0(fishing_efficiency_dat$year,"-01-01"))


ggplot(fishing_efficiency_dat, aes(year, y))+
  geom_line()+
  facet_grid(variable ~., scales = "free_y", labeller = label_wrap_gen())+
  scale_x_date()+
  labs(x="Year",y="Value",
       caption = paste(strwrap("Source: Fig 3 and Table 5from Greer et al 2019, https://doi.org/10.1016/j.marpol.2018.12.001",width = 50), collapse = "\n"))+
  theme_bw()


```



## Shipping

```{r load in iea total shipping emissions}
shipping_emissions <- readxl::read_excel(here::here("data/raw-data/external/IEA-Energy-End-uses-and-Efficiency-Indicators-database-Highlights-November-2023.xlsx"), sheet = "Transport - Emissions")

colnames(shipping_emissions) <- c("country","vehicle_type","product","2000","2005","2010","2015","2016","2017","2018","2019","2020","2021")

shipping_emissions <- shipping_emissions %>%
  filter(vehicle_type == "Total ships")

shipping_emissions <- reshape2::melt(shipping_emissions, 
                                     measure.vars = c("2000","2005","2010","2015","2016","2017","2018","2019","2020","2021"),
                                     variable.name = "year",
                                     value.name = "value")

shipping_emissions$year <- as.Date(paste0(shipping_emissions$year,"-01-01"))
shipping_emissions$value <- as.numeric(shipping_emissions$value)

shipping_emissions %>%
 filter(country == "IEA Total") %>%
  ggplot(aes(year, value))+
  geom_line()+
  geom_point()+
  scale_x_date()+
  labs(y = "IEA total shipping emissions (Mt CO2)")+
  theme_bw()+
  theme(legend.position = 'none')

```

```{r load in IEA shipping energy efficiency}

shipping_efficiency <- readxl::read_excel(here::here("data/raw-data/external/IEA-shipping-energy-efficiency.xlsx"), sheet = "Data")
shipping_efficiency$year <- as.Date(paste0(shipping_efficiency$year,"-01-01"))

ggplot(shipping_efficiency, aes(year, carbon_intensity_gCO2_per_tkm))+
  geom_line()+
  geom_point()+
  scale_x_date()+
  labs(y="Shipping energy efficiency (gCO2/tkm)", caption = "source: IEA energy end uses and efficiency indicators database")+
  theme_bw()
```


# Blue carbon

## Duarte et al 2020 BC Restoration projects

Source: Duarte, Carlos Manuel; Devassy, R P; Predragovic, M; Valuzzi, L; Parry, A; Hughes, Terry P (2020): Data set on restoration projects of coastal marine habitats reported worldwide [dataset]. PANGAEA, https://doi.org/10.1594/PANGAEA.912232. Accessed 24-04-2020

```{r}
bc_restoration <- readxl::read_excel(here::here("data/raw-data/external/Duarte2020RestorationDataSet-1.xlsx"),
                                     sheet = "Data")

bc_restoration %>%
  group_by(Habitat, Year) %>%
  summarise(n_projects = n()) %>%
  ungroup() %>% 
  tidyr::complete(Habitat, Year, fill = list(n_projects = 0)) %>%
  arrange(Habitat, Year) %>%
  group_by(Habitat) %>%
  mutate(year = as.Date(paste0(Year, "-01-01")),
         cumulative_projects = cumsum(n_projects)) %>%
  filter(Habitat != "Coral Reefs") %>%
  ggplot(aes(year, cumulative_projects, col = Habitat, fill = Habitat))+
  geom_area(position = "stack", alpha = 0.5)+
  labs(y = "Cumulative restoration projects (N)", x="Year",
       caption = "Source: Duarte et al (2020), doi: https://doi.org/10.1594/PANGAEA.912232")+
  scale_x_date()+
  theme_bw()+
  theme(legend.position = "bottom")

```















