---
title: "0.1_2025_OOSC_figures"
author: "Devi Veytia"
date: "2025-04-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Do just a sub-analysis for MRE at world-level


```{r load libraries}

# general data handing
library(dplyr)
library(dbplyr)
library(R.utils)
library(ggplot2)
library(ggalluvial)
library(tidyr)
library(stringr)
library(viridis)
library(countrycode)
library(broom)
library(conflicted)
library(tidyverse)

# bayesean analysis
library(rjags)


conflict_prefer("select", "dplyr")
conflicts_prefer(dplyr::filter)



## AESTHETICS
factor_aes <- readxl::read_excel(here::here("R/mitigation_factor_aesthetics.xlsx"))
typeAES <- factor_aes[which(factor_aes$variable == "oro_type"),]
typeAES <- typeAES[order(typeAES$order),]


```


# 1. Format data

Data structure:
list with each level is 
- data frame for publications: 
- data frame for n policy documents
- data frame for deployment

id variables: oro_type, component (publication, policy, deployment), variable_name,
response variable: y


## N publications

```{r read in oro type predictions v1}

# get id and year metadata
require(RSQLite)
p1_db <- RSQLite::dbConnect(RSQLite::SQLite(),
                      here::here("data","sqlite-databases","product1.sqlite"),
                                 create=FALSE)

uniquerefs <- tbl(p1_db, "uniquerefs") %>%
  select(analysis_id, year) %>%
  collect()

dbDisconnect(p1_db)

# read in predictions and join with year
predOroType <- read.csv(here::here("outputs/coded-label-predictions/oro_type_v1_predictions.csv")) %>%
  select(contains("mean_prediction"), id) %>%
  rename(analysis_id = id) %>%
  filter(if_any(-analysis_id, ~ . >= 0.5))%>%
  left_join(uniquerefs, by="analysis_id") %>%
  mutate(year = as.numeric(year)) %>%
  filter(!is.na(year))

colnames(predOroType) <- gsub("[.]","-", gsub("...mean_prediction|oro_type.","", colnames(predOroType)))

# how many relevant for each oro type?
apply(predOroType, 2, function(x) sum(0.5 <= x))

pubs <- predOroType %>%
  reshape2::melt(id.vars = c("analysis_id", "year"), variable.name = "oro_type", value.name = "mean") %>%
  filter(0.5 <= mean) %>%
  group_by(oro_type, year) %>%
  summarise(
    y = n_distinct(analysis_id)
  )%>%
  mutate(
    component = "publications",
    variable_name = "Publications (N)"
  )



ggplot(pubs, aes(x=year, y=y, alluvium=oro_type, fill=oro_type, colour=oro_type))+
  geom_alluvium(alpha = 0.8, decreasing=FALSE)+
  theme_minimal()

# save(pubs, file = here::here("data", "derived-data", "mitigationORO_pubs.RData"))
```

```{r Load in number of O&C publications to calculate proportion?}
## Load in predictions from previous chunk
load(here::here("data", "derived-data", "mitigationORO_pubs.RData"))
# Factor aesthetics
preds$year <- as.numeric(preds$year)
# preds$oro_type_f <- factor(
#  preds$oro_type, 
#   levels = typeAES$level, labels = typeAES$frenchLabel
# )


## Join with the number of O&C publications to get the proportion of ORO publications
wosDir <- here::here("data/external/ocean-and-climate-publications")
wosFiles <- dir(wosDir)

pubs <- data.frame()
for(g in 1:length(countryGroups)){
  for(c in 1:length(countryGroups[[g]])){
    
    # get OandA publications
    tempOA <- read.delim(paste(wosDir, wosFiles[grep(countryGroups[[g]][c], wosFiles)], sep="/")) %>%
    rename(
      year = Publication.Years,
      n_OC = Record.Count
    ) %>%
    select(year, n_OC) %>%
    mutate(year = as.numeric(year))

    
    # Tabulate number of publications
    cName = countries_ls$name_en[countries_ls$country == countryGroups[[g]][c]]
    tempPub <- preds[grep(cName, preds$countries),] %>%
        #left_join(docTotals, by = "analysis_id") %>%
        #filter(1980 <= year) %>% 
        group_by(oro_branch, oro_type, year) %>%
        summarise(n_ORO = n_distinct(analysis_id)) 
    
    # Join together and calculate proportion
    tempDf <- tempPub %>%
        left_join(tempOA, by = "year") %>%
        mutate(prop_ORO = n_ORO/n_OC,
               country_name = countryGroups[[g]][c],
               country_iso = countryGroupsIso[[g]][c]
               ) 
    
    # Bind all results
    pubs <- rbind(pubs, tempDf)
    
  }
  
}
rm(tempOA, tempPub, tempDf)
pubs <- pubs %>%
  filter(!is.na(year)&
           !is.na(n_ORO) &
           !is.na(prop_ORO))



## Do I calculate the weighted sum or count one document multiple times? For now comment out weighted sum calculations
# docTotals <- table(predMitFr$analysis_id) %>% as.data.frame(responseName = "doc_weight") 
# colnames(docTotals) <- c("analysis_id","doc_weight")
# docTotals[,2] <- 1/docTotals[,2]
# docTotals$analysis_id <- as.double(as.character(docTotals$analysis_id))



## Format for consistency with id variables:
# id variables: oro_branch, oro_type, country_name, country_iso, component (publication, policy, deployment), variable_name,
# response variable: y
pubs <- pubs %>%
  mutate(
    component = "publication"
  ) %>%
  reshape2::melt(id.vars = c("country_name","country_iso","component","oro_branch","oro_type","year"),
       value.name = "y", variable.name = "variable_name")




# save(pubs, file = here::here("data", "derived-data", "prop_ORO_pubs.RData"))
```






## N policy documents

```{r read in legislation documents}

## Read in and format policy matches

legDat <- read.csv(here::here("data/derived-data/en-fr-de-sp_fullTextQueryMatches_allOROs_long.csv")) %>%
  rename(country_name = Country_short, year=Year, oro_type = oroType) %>%
  filter(1975 < year,
         Binding.legislation == "True", 
         National.level == "True"
         ) %>%
  mutate(
    oro_type = case_when(
      oro_type == "M_MRE_general" ~ "MRE-Located", ## Note to fix this later
      oro_type == "M_MRE_ocean" ~ "MRE-Ocean",
      oro_type == "M_MRE_located" ~ "MRE-Located",
      oro_type == "M_MRE_bio" ~ "MRE-Bio",
      oro_type == "M_Incr_eff" ~ "Efficiency",
      oro_type == "M_CCS" ~ "CCS",
      oro_type == "M_CDR_bc" ~ "CDR-BC",
      oro_type == "M_CDR_oae" ~ "CDR-OAE",
      oro_type == "M_CDR_biopump" ~ "CDR-BioPump",
      oro_type == "M_CDR_cult" ~ "CDR-Cult",
     
      TRUE ~ "Other"
    )) %>%
  filter(oro_type != "Other") %>%
  # Tabulate all documents
  group_by(oro_type, year) %>%
  summarise(y = n_distinct(search_id)) %>%
  mutate(
    component = "legislation",
    variable_name = "Legislation (N)"
  )




## quick plot
  
pol_stackedBP <- ggplot(legDat, aes(x=year, y=y, fill=oro_type))+
  geom_col(position="stack")+
  scale_fill_manual(name = "ORO type",values = typeAES$colour, breaks = typeAES$level)+
  theme_minimal()+
  theme(legend.position = "bottom")
pol_stackedBP

## save
save(legDat, file = here::here("data", "derived-data", "n_legislation_docs.RData"))

```

```{r read in non-binding policy documents}

## Read in and format policy matches

polDat <- read.csv(here::here("data/derived-data/en-fr-de-sp_fullTextQueryMatches_allOROs_long.csv")) %>%
  rename(country_name = Country_short, year=Year, oro_type = oroType) %>%
  filter(1975 < year,
         Binding.legislation == "False", 
         National.level == "True"
         ) %>%
  mutate(
    oro_type = case_when(
      oro_type == "M_MRE_general" ~ "MRE-Located", ## Note to fix this later
      oro_type == "M_MRE_ocean" ~ "MRE-Ocean",
      oro_type == "M_MRE_located" ~ "MRE-Located",
      oro_type == "M_MRE_bio" ~ "MRE-Bio",
      oro_type == "M_Incr_eff" ~ "Efficiency",
      oro_type == "M_CCS" ~ "CCS",
      oro_type == "M_CDR_bc" ~ "CDR-BC",
      oro_type == "M_CDR_oae" ~ "CDR-OAE",
      oro_type == "M_CDR_biopump" ~ "CDR-BioPump",
      oro_type == "M_CDR_cult" ~ "CDR-Cult",
     
      TRUE ~ "Other"
    )) %>%
  filter(oro_type != "Other") %>%
  # Tabulate all documents
  group_by(oro_type, year) %>%
  summarise(y = n_distinct(search_id)) %>%
  mutate(
    component = "non-binding policy",
    variable_name = "Non-binding policy (N)"
  )




## quick plot
  
pol_stackedBP <- ggplot(polDat, aes(x=year, y=y, fill=oro_type))+
  geom_col(position="stack")+
  scale_fill_manual(name = "ORO type",values = typeAES$colour, breaks = typeAES$level)+
  theme_minimal()+
  theme(legend.position = "bottom")
pol_stackedBP

## save
save(polDat, file = here::here("data", "derived-data", "n_nonBindPolicy_docs.RData"))

```



## N deployments




Difficult to get a dataset with enough longevity, so could combine indicators?

N patents from altmetrics (long, country and year available)


field trials (time coverage ~ 2018 - 2023, spatially resolved)
- (Ocean visions field trail database)[]

mCDR companies/ start-ups : goal: number of startups by founding year (recent. Need to get country and year)
-  https://community.oceanvisions.org/organizations - company names
- https://www.cdr.fyi/leaderboards - company names, tons sold, tons delivered
- GESAMP The current proposals on ocean climate intervention that GESAMP is aware of as of 31st May 2024: GESAMP_wg41_ocean_climate_intervention_projects_31_may_2024.xlsx http://www.gesamp.org/work/groups/41 ~ 184 company names
- OceanNETs_D18_oceanbased_CDR_companies ~ 50 companies with country by no year


I think for now, stick with patents and field trials, and maybe OceanNET CDR companies



```{r MRE deployment - Installed capacity}
## load data
# Data source: IRENA renewable electricity statistics (https://www.irena.org/Publications/2024/Jul/Renewable-energy-statistics-2024)

MreDat <- readxl::read_excel(here::here("data/raw-data/external/IRENA-electricity-statistics-by-country-year-2024.xlsx"),
                         sheet = "Country") %>% 
  filter(
    Technology %in% c("Marine energy","Offshore wind energy")
  ) %>% 
  mutate(
    oro_type = case_when(
      Technology == "Marine energy" ~ "MRE-Ocean",
      Technology == "Offshore wind energy" ~ "MRE-Located"
    )
  ) %>%
  rename(year = Year)  %>%
  group_by(year,oro_type) %>% 
  summarise(y = sum(`Electricity Installed Capacity (MW)`, na.rm=T)) %>%
  # add identification columns
  mutate(
    year = as.numeric(year),
    component = "deployment",
    variable_name = "Electricity Installed Capacity (MW)"
    
  )

ggplot(MreDat, aes(x=year, y=y, fill=oro_type))+
  geom_col(position="stack")+
  scale_fill_manual(name = "ORO type",values = typeAES$colour, breaks = typeAES$level)+
  theme_minimal()+
  theme(legend.position = "bottom")


```


```{r Increase efficiency - domestic freight energy efficiency}
# IEA domestic freight transport energy efficiency (gCO2/tkm)
# https://www.iea.org/data-and-statistics/data-product/energy-efficiency-indicators
eff <- readxl::read_excel(
  here::here("data/raw-data/external/IEA-shipping-energy-efficiency.xlsx"), 
  sheet = "Data") %>%
  mutate(
    y= 1/carbon_intensity_gCO2_per_tkm,
    year = as.numeric(format(as.Date(paste0(year,"-01-01")), "%Y")),
    variable_name = "Shipping carbon efficiency (tkm/gCO2)",
    component = "deployment",
    oro_type = "Efficiency"
    ) 

ggplot(eff, aes(x=year, y=y, fill=oro_type))+
  geom_col()+
  scale_fill_manual(name = "ORO type",values = typeAES$colour, breaks = typeAES$level)+
  theme_minimal()+
  theme(legend.position = "bottom")


```

For CCS download from [oil and gas climate initiative](https://www.ogci.com/ccus/co2-storage-catalogue). Choose to sum the variable of Storage Capacity: Those quantities of Total Storage Resources anticipated to be commercially accessible in the characterized geologic formation by application of development projects from a given date forward under defined conditions. Commercial Storage Resources must further satisfy four criteria: The target geologic formation must be discovered and characterized (including containment); it must be possible to inject at the required rates; the development project must be commercial; and the storage resource must remain, as of the evaluation date (i.e., not previously used for storage), on the development project(s) applied (https://www.ogci.com/ccus/co2-storage-catalogue/co2-storage-catalogue-about-the-catalogue) -- but only if associated with a project 

```{r CCS}


ccsDat <- readxl::read_excel(here::here("data/raw-data/external/CRSC_CYCLE_4_FINAL_2024_160724.xlsx")) %>%
  filter(grepl("shore|sea", area, ignore.case=T), !is.na(year_of_publication)) %>%
  filter(project_spec == "YES") %>%
  rename(year = year_of_publication) %>%
  group_by(year) %>%
  summarise(
    y = sum(sum_low, sum_mid, sum_high, na.rm=T) #sum(stored_low, stored_mid, stored_high, na.rm=T)
  ) %>%
  mutate(
    year = as.numeric(year),
    variable_name = "Storage capacity (Mt)",
    component = "deployment",
    oro_type = "CCS"
  )

ggplot(ccsDat, aes(year, y))+
  geom_col()

```


Maybe can relate blue carbon to BC variable crossed with restoration

```{r CDR-BC - number of restoration projects Duarte et al 2020}
bc_restoration <- readxl::read_excel(here::here("data/raw-data/external/Duarte2020RestorationDataSet-1.xlsx"),
                                     sheet = "Data")%>%
  filter(Habitat %in% c("Mangrove","Saltmarshes","Seagrass meadows")) %>% # Omit Oyster Reefs and Coral Reefs
  # Sum the number of projects /year
  mutate(year = as.numeric(Year)) %>%
  group_by(year) %>%
  summarise(
    y = n()
  ) %>%
  ungroup() %>% 
  # # Calculate cumulative value?
  # tidyr::complete(year, fill = list(y = 0)) %>%
  # arrange(year) %>%
  # mutate(y = cumsum(y)) %>%
  mutate(
    variable_name = "BC restoration projects (N)",
    component = "deployment",
    oro_type = "CDR-BC"
  )

head(bc_restoration)

ggplot(bc_restoration, aes(x=year, y=y))+
  geom_col()
```

Other CDR not available at country level. 
Make proportional to amount of available habitat area? https://essd.copernicus.org/articles/16/3433/2024/#&gid=1&pid=1 -- van Zelst et al 2022

```{r CDR-OAE CDR-Cult CDR-BioPump}
# Ocean visions field trials 
fieldTrials <- readxl::read_excel(
  here::here("data/raw-data/external/Ocean-visions-mCDR-field-trial-database.xlsx"),
  sheet = "Data") %>%
  mutate(
    `Start of Pilot` = replace(`Start of Pilot`, `Start of Pilot` %in% c("5.12.2023","2023"), as.numeric(as.Date("2023-01-01"))),
    trialID = row_number()
  ) %>%
  mutate(year = as.Date(as.numeric(`Start of Pilot`), origin = "1899-12-30")) %>%
  mutate(year = ifelse(year == as.Date("1952-12-30"), as.Date("2023-01-01"), year)) %>%
  mutate(year = as.Date(year, origin = as.Date("1970-01-01"))) %>%
  separate_rows(`All CDR Methods`, sep=",") %>% 
  mutate(
    oro_type = case_when(
      grepl("OAE|Alkalinity|Weathering", `All CDR Methods`) ~ "CDR-OAE",
      grepl("Biomass|Macroalgae|Cultivation", `All CDR Methods`) ~ "CDR-Cult",
      grepl("Upwelling", `All CDR Methods`) ~ "CDR-BioPump",
      TRUE ~ "Other"
    ),
    year = as.numeric(format(year, "%Y")),
    source = "Field trial"
  ) %>%
  filter(!is.na(year) & oro_type != "Other") 
  # group_by(oro_type, year) %>%
  # summarise(
  #   y=n_distinct(trialID)
  # ) %>%
  # mutate(
  #   variable_name = "Field Trials (N)",
  #   component = "deployment"
  # )


# Founding year of mcdr startups
# Need to de-duplicate and add year  

GESAMP_companies <- readxl::read_excel(
  here::here("data/raw-data/external/GESAMP_wg41_ocean_climate_intervention_projects_31_may_2024.xlsx"), sheet="AllTables") %>%
  select(Company, Type, Website) %>%
  mutate(
    Company = trimws(gsub("[[:punct:]]", "", Company)),
    source = "GESAMP WG 41"
  )%>%
  filter(!is.na(Company))

# Load in OceanNET companies and de-deuplicate from GESAMP companies
OceanNET_companies <- read.csv(
  here::here("data/raw-data/external/OceanNETs_D18_oceanbased_CDR_companies/dataset/D1_8_database_oceanbased_companies.csv")) %>%
  mutate(
    Company = trimws(gsub("[[:punct:]]", "", Company)),
    source = "OceanNETs (2020) D1.8 database"
  )%>%
  filter(!is.na(Company), !is.na(Lat))%>%
  select(-c(contains("X"))) %>%
  filter(!(tolower(Company) %in% tolower(GESAMP_companies$Company)), Company != "Qilibrium")


# Join
mCDRCompanies <- GESAMP_companies %>%
  bind_rows(OceanNET_companies)

# write to file
write.csv(mCDRCompanies, file = here::here("data/derived-data/mCDR-companies-dedup.csv"))

## read in the years
mCDRCompanies <- readxl::read_excel(here::here("data/raw-data/mCDR-companies-linkedin.xlsx")) %>%
  filter(!is.na(`Company founded on (linkedin)`)) %>%
  rename(year = `Company founded on (linkedin)`) %>%
  select(Company, year) %>%
  left_join(GESAMP_companies %>% select(Company, Type), by="Company") %>%
  left_join(OceanNET_companies %>% select(Company, Type), by="Company") %>%
  mutate(
    Type = ifelse(is.na(Type.x), Type.y, Type.x)
  ) %>%
  select(Company, Type, year) %>%
  mutate(
    oro_type = case_when(
      grepl("Upwell|fertilization", Type, ignore.case=TRUE) ~ "CDR-BioPump",
      grepl("Biomass sinking|Farming|Harvesting|Aquaculture", Type, ignore.case=TRUE) ~ "CDR-Cult",
      grepl("OAE|Alkalinity|weathering", Type) ~ "CDR-OAE",
      TRUE~"CDR-Other"
    ),
    source = "Startup",
    year = as.numeric(year)
  ) %>%
  filter(oro_type != "CDR-Other")
  
# Ocean visions community and www.cdr.fyi/leaderboards can't be scraped because they use private APIs
# Although the latter a list of suppliers can be found here without metadata
# https://www.cdr.fyi/api/search




## Combine data sources
mCDRDeploy <- fieldTrials[,c("oro_type","year","source")] %>%
  bind_rows(mCDRCompanies[,c("oro_type","year","source")]) %>%
  group_by(oro_type, year) %>%
  summarise(
    y=n()
  ) %>%
  mutate(
    variable_name = "Field trial/startup company (N)",
    component = "deployment"
  )

ggplot(mCDRDeploy, aes(year, y, fill=oro_type))+
  geom_col()

```




```{r save all deployment data, eval=FALSE}
cols <- c("component","oro_type","year","variable_name","y")

allDeployDat <- rbind(
  MreDat[,cols],
  eff[,cols],
  ccsDat[,cols],
  bc_restoration[,cols],
  mCDRDeploy[,cols]
)

ggplot(allDeployDat, aes(year, y, col=oro_type))+
  geom_line()+
  facet_wrap(vars(oro_type), scales="free")+
  geom_text(aes(label = str_wrap(variable_name, 25)), x=-Inf, y=Inf, check_overlap = T, guide=FALSE, vjust=1, hjust=0, size=3)+
  theme(
    legend.position = "none"
  )

ggsave(here::here("figures/supplemental/mitigationDeploymentIndicators.pdf"), width = 15, height=12)

save(allDeployDat, file=here::here("data/derived-data/mitigationDeployDat.RData"))

```


# Visualize

```{r plot multiplanel alluvial}

# scale deployment data
allDeployDat_scale <- allDeployDat %>%
  filter(!is.na(year)) %>%
  arrange(oro_type, year) %>%
  group_by(oro_type) %>%
  mutate(
    y=cut(
      scales::rescale(y, to=c(0,1)), 
      breaks = seq(0,1, by = 0.1), 
      labels = seq(0.1,1, by = 0.1)
      ) %>% as.numeric()# if i do 0,1 it throws an error
  ) %>%
  ungroup() 
  

## plot all together

allComponentDat_scale <- pubs %>%
  bind_rows(polDat) %>%
  bind_rows(legDat) %>%
  bind_rows(allDeployDat_scale) %>%
  mutate(
    component = factor(component, 
                       levels = c("publications","non-binding policy","legislation","deployment"), 
                       labels = c("Academic\n(N publications)", 
                                  "Non-binding policy\n(N documents)", 
                                  "Legislation\n(N documents)",
                                  "Action\n(scaled metric)"))
  )



allComponentAlluvial <- ggplot(allComponentDat_scale %>% filter(1980 <= year, year <= 2022), aes(x=year, y=y, alluvium=oro_type, fill=oro_type, colour=oro_type))+
  # geom_alluvium(alpha = 0.8, decreasing=FALSE)+
  geom_alluvium(alpha = 0.8, curve_type = "sigmoid", decreasing=FALSE) +
  labs(y="",x="Year")+
  facet_grid(component~., scales = "free_y", switch="y")+
  scale_fill_manual(name = "ORO type",values = typeAES$colour, breaks = typeAES$level, labels = typeAES$label)+
  scale_colour_manual(name = "ORO type",values = typeAES$colour, breaks = typeAES$level, labels = typeAES$label)+
  # scale_x_continuous(limits = c(1980,2022))+
  guides(color = guide_legend(nrow = 3), fill = guide_legend(nrow = 3))+
  theme_bw()+
  theme(
    legend.position = "bottom", 
    strip.placement = "outside",
    legend.title = element_blank()
    )


## try on log scale?
allComponentAlluvial_log <- ggplot(allComponentDat_scale %>% 
                                     mutate(y=log(y)) %>% 
                                     filter(1990 <= year, year <= 2022) %>%
                                     na.omit() %>% 
                                     mutate(
                                       y= cut(y,
                                              breaks = seq(0,8, by = 0.1), 
                                              labels = seq(0.1,8, by = 0.1)
                                              ) %>% as.numeric()
                                     ), 
                                   aes(x=year, y=y, alluvium=oro_type, fill=oro_type, colour=oro_type))+
  # geom_alluvium(alpha = 0.8, decreasing=FALSE)+
  geom_alluvium(alpha = 0.8, curve_type = "sigmoid", decreasing=FALSE) +
  labs(y="",x="Year", caption = "y axis on log scale")+
  facet_grid(component~., scales = "free_y", switch="y")+
  scale_fill_manual(name = "ORO type",values = typeAES$colour, breaks = typeAES$level, labels = typeAES$label)+
  scale_colour_manual(name = "ORO type",values = typeAES$colour, breaks = typeAES$level, labels = typeAES$label)+
  # scale_x_continuous(limits = c(1980,2022))+
  # guides(color = guide_legend(nrow = 3), fill = guide_legend(nrow = 3))+
  guides(color = guide_legend(ncol = 1), fill = guide_legend(ncol = 1))+
  theme_bw()+
  theme(
    legend.position = "right", 
    strip.placement = "outside",
    legend.title = element_blank()
    )


ggsave(
  plot = allComponentAlluvial, 
  filename = here::here("figures/2025-06-02_OOSC/allComponentAlluvial.png"),
  width=7, height=6, units="in", dpi = 600
    )


ggsave(
  plot = allComponentAlluvial_log, 
  filename = here::here("figures/2025-06-02_OOSC/allComponentAlluvial_log.png"),
  width=7, height=6, units="in", dpi = 600
    )
```

# Analyses

## SEM

https://stats.oarc.ucla.edu/r/seminars/rsem/

Maybe can't resolve at the ORO level and need to pool

```{r explore SEM}

library(vars)
library(lavaan)
library(semPlot)
library(tidyr)

lag_ap <- c(1,2)
lag_pl <- c(1)
lag_la <- c(1)
lag_ap <- c(1)
lags_combinations <- expand.grid(lag_ap, lag_pl, lag_la,lag_ap)
year_lim <- c(2000, 2022)


# Format data, and take log -- 
allComponentDat_model <- pubs %>%
  bind_rows(polDat) %>%
  bind_rows(legDat) %>%
  bind_rows(allDeployDat_scale) %>%
  mutate(
    component = replace(component, component == "non-binding policy", "policy"),
    # y = scale(y, center=TRUE, scale=TRUE)
  ) %>%
  filter(year_lim[1] <= year & year <= year_lim[2])

ggplot(allComponentDat_model, aes(x=year, y=y, fill = oro_type))+
  geom_col()+
  facet_wrap(component ~., scales = "free_y")

# Split by oro type
oro_groups <- split(allComponentDat_model, allComponentDat_model$oro_type)

# Function to fit model

# variables c("publications","non-binding policy","legislation","deployment")

model <- '
    # All directional hypotheses
    
    policy_lag ~ 1 + publications
    legislation_lag ~ 1+ policy_lag 
    action_lag ~ 1 + legislation_lag
    publications_lag ~ 1+action_lag
  '



analyze_group <- function(df_group, oro_label) {
  
  # For testing
  df_group <- oro_groups[[1]]
  oro_label <- names(oro_groups)[1]
  
  # scale?
  df_group <- df_group %>%
    dplyr::select(year, component, y) %>%
    group_by(component) %>%
    mutate(y = scale(y, center=TRUE))
  ggplot(df_group, aes(year, y))+ geom_col()+facet_wrap(component ~., scales="free_y")
  
  # Reshape to wide format, replace NA with 0
  df_wide <- df_group %>%
    pivot_wider(names_from = component, values_from = y) %>%
    arrange(year) %>%
    mutate(across(-c(year), ~ replace_na(., 0))) 
  
  # Calculate lagged variables
  lag_ap <- c(1,2) # academic to policy
  lag_pl <- c(1) # policy to legislation
  lag_ld <- c(1) # legislation to deployment
  lag_da <- c(1) # deployment to academic
  lags_combinations <- expand.grid(lag_ap, lag_pl, lag_ld,lag_da)
  colnames(lags_combinations) <- c("lag_ap", "lag_pl", "lag_ld","lag_da")
  lags <- lags_combinations[1,] # for testing
  
  
  model <- '
    policy_lag ~ 1 + publications
    legislation_lag ~ 1+ policy_lag 
    deployment_lag ~ 1 + legislation_lag
    publications_lag ~ 1+deployment_lag
  '
  
  df_wide <- df_wide %>%
    mutate(
      policy_lag = lag(policy, sum(lags[1])),
      legislation_lag = lag(legislation, sum(lags[1:2])),
      deployment_lag = lag(deployment, sum(lags[1:3])),
      publications_lag = lag(publications, sum(lags[1:4]))
    ) %>%
    dplyr::select(year, publications, contains("lag"))%>%
    na.omit()
  

  
  # Fit a simple SEM: test all directions
  fit <- sem(model, data = df_wide, missing = "ML")

  # Visualize SEM results
  semPlot::semPaths(fit, whatLabels = "est", layout = "circle")

  # Return results for further processing
  list(change_points = change_points, ccf_matrix = ccf_matrix, sem_fit = fit)
}

# Run analysis for each group
results <- lapply(names(oro_groups), function(name) {
  analyze_group(oro_groups[[name]], name)
})
names(results) <- names(oro_groups)

```


```{r}
# Load required packages

library(changepoint)
library(lmtest)
library(vars)
library(lavaan)
library(semPlot)
library(tidyr)

# Load the data
data <- read.csv(here::here("data/derived-data/mitigationDeployDatTable.csv"))

# Ensure time is in correct format
data$time <- as.numeric(as.Date(paste0(data$year, "-01-01")))

# Split by 'oro_type'
oro_groups <- split(data, data$oro_type)

# Function to perform analysis per oro_type
analyze_group <- function(df_group, oro_label) {
  # Reshape to wide format
  df_wide <- df_group %>%
    select(time, component, y) %>%
    pivot_wider(names_from = component, values_from = y) %>%
    arrange(time)

  # --- Change point detection for each component ---
  change_points <- lapply(df_wide[-1], function(series) {
    tryCatch({
      cpt.meanvar(series, method = "PELT", penalty = "SIC")$cpts
    }, error = function(e) NA)
  })

  print(paste("Change points for", oro_label))
  print(change_points)

  # --- Cross-correlation analysis ---
  # Create time series
  ts_data <- ts(df_wide[-1], start = min(df_wide$time), frequency = 1)
  ccf_matrix <- list()

  comps <- colnames(ts_data)
  for (i in 1:length(comps)) {
    for (j in 1:length(comps)) {
      if (i != j) {
        ccf_res <- ccf(ts_data[, i], ts_data[, j], plot = FALSE)
        max_corr <- max(abs(ccf_res$acf))
        lag_at_max <- ccf_res$lag[which.max(abs(ccf_res$acf))]
        ccf_matrix[[paste(comps[i], comps[j], sep = "_")]] <- c(max_corr, lag_at_max)
      }
    }
  }
  print(paste("Cross-correlation results for", oro_label))
  print(ccf_matrix)

  # --- Structural Equation Modeling (SEM) ---
  # Fit a simple SEM: test all directions
  sem_data <- na.omit(df_wide[-1])  # exclude time, and drop NAs
  model <- '
    # All directional hypotheses
    publications ~ policy + deployment
    policy ~ publications + deployment
    deployment ~ publications + policy
  '
  fit <- sem(model, data = sem_data, missing = "ML")

  # Visualize SEM results
  semPlot::semPaths(fit, whatLabels = "est", layout = "circle", title = paste("SEM for", oro_label))

  # Return results for further processing
  list(change_points = change_points, ccf_matrix = ccf_matrix, sem_fit = fit)
}

# Run analysis for each group
results <- lapply(names(oro_groups), function(name) {
  analyze_group(oro_groups[[name]], name)
})
names(results) <- names(oro_groups)


```






